--- 
title: "Theory Construction and Statistical Modeling"
#subtitle: "A guide to structural equation modeling in R"
author: "Kyle M. Lang"
date: "Last updated: `r Sys.Date()`"
site: bookdown::bookdown_site
#output: 
#  bookdown::gitbook:
#    config: 
#      toc:
#       collapse: section
#      search: yes
#      fontsettings:
#        size: 2
#    split_by: section
#    df_print: paged
#documentclass: book
#bibliography: [../bibtex/book.bib, ../bibtex/packages.bib]
#biblio-style: apalike
#link-citations: yes
#description: "This is a guide for the elective course TCSM."
---

# Course Information {-}

```{r global_options, include = FALSE}
library(knitr)
library(dplyr)

opts_chunk$set(warning = FALSE, message = FALSE)

dataDir  <- "../../../data/"
schedule <- read.csv2("../data/schedule.csv")
```

In order to test a theory, we must express the theory as a statistical model and 
then test this model on quantitative (numeric) data. This course uses existing 
tutorial datasets to practice the process of translating verbal theories into 
testable statistical models. If you are interested in the methods of acquiring 
high quality data to test your own theory, we recommend following the course 
*Conducting a Survey* which is taught from November to January.

In this course we will use datasets from different disciplines within the social 
sciences (educational sciences, psychology, and sociology) to explain and 
illustrate theories and practices that are used in all social science disciplines to statistically model social science theories.

Most information about the course is available in this GitBook. Course-related 
communication will be through <https://uu.blackboard.nl> (Log in with your 
student ID and password).

## Instructors

**Coordinator:**

[dr. Kyle M. Lang](https://www.kylemlang.com)

**Lectures**

[dr. Kyle M. Lang](https://www.kylemlang.com)

**Practicals**

[Rianne Kraakman](https://www.uu.nl/staff/amkraakman)

[Daniëlle Remmerswaal](https://www.uu.nl/staff/DMRemmerswaal)

## Course overview

This course comprises three parts:

1. *Factor analysis*: You will learn different ways of defining and estimating
unobserved constructs.
1. *Path analysis*: You will learn how to conduct regressions and ANOVAs as 
structural equation models with observed variables.
1. *Full structural equation modeling*: You will combine the first two topics by 
estimating path models using latent variables.

Each of these three themes will be evaluated with a separate assignment. Your 
course grade will be based on the weighted average of these three assignment 
grades.

### Schedule

```{r print_schedule, echo = FALSE}
kable(schedule, 
      col.names = gsub("\\.", " ", colnames(schedule))
      )
```

```{r, results = "asis", echo = FALSE, eval = FALSE}
library(DT)
# library(lubridate)
# mdy(isoweek(4))
# lubridate::isoweek(as.Date("13-09-2021", format = "%d-%m-%Y"))
# lubridate::parse_date_time("37", orders = "%W")
# lubridate::isoweek(as.Date("13-09-2021", format = "%d-%m-%Y"))
schedule$Date <- as.Date(schedule$Date, format = "%d-%m-%Y")
library(DT)
DT::datatable(schedule[order(schedule$Date, decreasing = FALSE), ], rownames = FALSE, options = list(pageLength = nrow(schedule))) |>
    formatStyle(
    'time',
    backgroundColor = styleEqual("14:15 - 15:00", 'yellow'))
```

## Learning goals

In this course you will learn how to translate a social scientific theory into a 
statistical model, how to analyze your data with these models, and how to 
interpret and report your results following APA standards.

After completing the course, you will be able to:

1. Translate a verbal theory into a conceptual model, and translate a conceptual 
model into a statistical model.
1. Independently analyze data using the free, open-source statistical software R.
1. Apply a *latent variable model* to a real-life problem wherein the observed 
variables are only indirect indicators of an unobserved construct.
1. Use a *path model* to represent the hypothesized causal relations among
several variables, including relationships such as mediation and moderation.
1. Explain to a fellow student how *structural equation modeling* combines 
latent variable models with path models and the benefits of doing so.
1. Reflect critically on the decisions involved in defining and estimating
structural equation models.

## Resources

### Literature

**You do not need a separate book for this course!** 

Most of the information is contained within this GitBook and the course readings 
(which you will be able to access via links in this GitBook). 

All literature is freely available online, as long as you are logging in from 
within the UU-domain (i.e., from the UU campus or through an appropriate VPN). 
All readings are linked in this GitBook via their DOIs. 

If you run into any trouble accessing a given article, searching for the title 
using [Google Scholar](www.scholar.google.com) or the 
[University Library](www.ubu.uu.nl) will probably due the trick.

### Software

You will do all of your statistical analyses with the statistical programming language/environment [R][r] and the add-on package [`lavaan`][lavaan].

- If you want to expand your learning, you can follow this excellent 
[`lavaan` tutorial](https://lavaan.ugent.be/tutorial/). Doing so is entirely 
optional, though.

## Reading questions

Along with every article, we will provide reading questions. You will not be 
graded on the reading questions, but it is important to prepare the reading 
questions before every lecture. The reading questions serve several important 
purposes:

-	Provide relevant background knowledge for the lecture
-	Help you recognize and understand the key terms and concepts
-	Make you aware of important publications that shaped the field
- Help you extract the relevant insights from the literature

## Weekly preparation

Before every class meeting (both lectures and practicals) you need to do the 
assigned homework (delineated in the GitBook chapter for that week). This course 
follows a *flipped classroom* procedure, so you must complete the weekly homework 
to meaningfully participate in, and benefit from, the class meetings.

### Background knowledge

We assume you have basic knowledge about multivariate statistics before entering 
this course. You do not need any prior experience working with R. If you wish to 
refresh your knowledge, we recommend the chapters on ANOVA, multiple regression,
and exploratory factor analysis from Field’s [*Discovering Statistics using R*][field]. 

If you cannot access the Field book, many other introductory statistics textbooks 
cover these topics equally well. So, use whatever you have lying around from 
past statistics courses. You could also try one of the following open-access options:

- [*Applied Statistics with R*][stats_with_r]
- [*Introduction to Modern Statistics*][modern_stats]
- [*Introduction to Statistical Learning*][isl]

## Grading

Your grade for the course is based on a “portfolio” composed of the three 
take-home assignments:

1. Latent variable modeling 

- Deadline: XXXX
- Group assignment
- Contributes 25% of your course grade

1. Path modeling 

- Deadline: XXXX
- Group assignment
- Contributes 25% of your course grade

1. Full Structural equation modeling 

- Deadline: XXXX
- Individual assignment
- Contributes 50% of your course grade

The specifics of the assignments are described in Chapter \@ref(assignments)

## Attendance

Attendance is not mandatory, but we strongly encourage you to attend all lectures 
and practicals. In our experience, students who actively participate tend to pass 
the course, whereas those who do not participate tend to drop out or fail. The 
lectures and practicals build on each other, so, in the unfortunate event that 
you have to miss a class meeting, please make sure you have caught up with the 
material before the next session.

[r]: https://www.r-project.org/
[lavaan]: https://www.lavaan.ugent.be/
[field]: https://www.discoveringstatistics.com/books/discovering-statistics-using-r/
[stats_with_r]: https://book.stat420.org/
[modern_stats]: https://openintro-ims.netlify.app/
[isl]: https://www.statlearning.com
