--- 
title: "Theory Construction and Statistical Modeling"
#subtitle: "A guide to structural equation modeling in R"
author: "Kyle M. Lang"
date: "Last updated: `r Sys.Date()`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    config: 
      toc:
       collapse: section
       search: yes
    split_by: section
    df_print: paged
documentclass: book
#bibliography: [../bibtex/book.bib, ../bibtex/packages.bib]
#biblio-style: apalike
#link-citations: yes
#description: "This is a guide for the elective course TCSM."
---

```{r global_options, include = FALSE}
library(knitr)
library(dplyr)

opts_chunk$set(warning = FALSE, message = FALSE)

dataDir  <- "../../data/"
```

# Course Information {-}

In order to test a theory, we must express the theory as a statistical model and 
then test this model on quantitative (numeric) data. In this course we will use 
datasets from different disciplines within the social sciences (educational 
sciences, psychology, and sociology) to explain and illustrate theories and 
practices that are used in all social science disciplines to statistically model 
social science theories.

This course uses existing tutorial datasets to practice the process of 
translating verbal theories into testable statistical models. If you are 
interested in the methods of acquiring high quality data to test your own theory, 
we recommend following the course *Conducting a Survey* which is taught from 
November to January.

Most information about the course is available in this GitBook. Course-related 
communication will be through <https://uu.blackboard.com> (Log in with your 
student ID and password).

<!----------------------------------------------------------------------------->

## Acknowledgement {-}

This course was originally developed by 
[dr. Caspar van Lissa](https://www.tilburguniversity.edu/staff/c-j-vanlissa). 
Indeed, you will still see Caspar in the lecture recordings. I 
([dr. Kyle M. Lang](https://www.kylemlang.com)) have modified Caspar's original 
materials and take full responsibility for any errors or inaccuracies introduced 
through these modifications. Credit for any particularly effective piece of 
pedagogy should probably go to Caspar.

- You can view the original version of this course 
[here](https://github.com/cjvanlissa/TCSM) on Caspar's GitHub page.

<!----------------------------------------------------------------------------->

## Instructors {-}

**Coordinator:**

[dr. Kyle M. Lang](https://www.kylemlang.com)

**Lectures:**

[dr. Kyle M. Lang](https://www.kylemlang.com)

**Practicals:**

[Rianne Kraakman](https://www.uu.nl/staff/amkraakman)

[Daniëlle Remmerswaal](https://www.uu.nl/staff/DMRemmerswaal)

Laura Jochim

<!----------------------------------------------------------------------------->

## Course overview {-}

This course comprises three parts:

1. *Factor analysis*: You will learn different ways of defining and estimating
unobserved constructs.
1. *Path analysis*: You will learn how to conduct regressions and ANOVAs as 
structural equation models with observed variables.
1. *Full structural equation modeling*: You will combine the first two topics by 
estimating path models using latent variables.

Each of these three themes will be evaluated with a separate assignment. Your 
course grade will be based on the weighted average of these three assignment 
grades.

### Schedule {-}

```{r print_schedule, echo = FALSE}
schedule <- read.csv2("data/schedule.csv")
knitr::kable(schedule, 
             col.names = gsub("\\.", " ", colnames(schedule))
             )
```

<!----------------------------------------------------------------------------->

## Learning goals {-}

In this course you will learn how to translate a social scientific theory into a 
statistical model, how to analyze your data with these models, and how to 
interpret and report your results following APA standards.

After completing the course, you will be able to:

1. Translate a verbal theory into a conceptual model, and translate a conceptual 
model into a statistical model.
1. Independently analyze data using the free, open-source statistical software R.
1. Apply a *latent variable model* to a real-life problem wherein the observed 
variables are only indirect indicators of an unobserved construct.
1. Use a *path model* to represent the hypothesized causal relations among
several variables, including relationships such as mediation and moderation.
1. Explain to a fellow student how *structural equation modeling* combines 
latent variable models with path models and the benefits of doing so.
1. Reflect critically on the decisions involved in defining and estimating
structural equation models.

<!----------------------------------------------------------------------------->

## Resources {-}

---

### Literature {-}

**You do not need a separate book for this course!** 

Most of the information is contained within this GitBook and the course readings 
(which you will be able to access via links in this GitBook). 

All literature is freely available online, as long as you are logging in from 
within the UU-domain (i.e., from the UU campus or through an appropriate VPN). 
All readings are linked in this GitBook via either direct download links or DOIs. 

If you run into any trouble accessing a given article, searching for the title 
using [Google Scholar][gs] or the [University Library][uu_lib] will probably due 
the trick.

---

### Software {-}

You will do all of your statistical analyses with the statistical programming 
language/environment [R][r] and the add-on package [`lavaan`][lavaan].

- If you want to expand your learning, you can follow this excellent 
[`lavaan` tutorial](https://lavaan.ugent.be/tutorial/). Doing so is entirely 
optional, though.

<!----------------------------------------------------------------------------->

## Reading questions {-}

Along with every article, we will provide reading questions. You will not be 
graded on the reading questions, but it is important to prepare the reading 
questions before every lecture. The reading questions serve several important 
purposes:

-	Provide relevant background knowledge for the lecture
-	Help you recognize and understand the key terms and concepts
-	Make you aware of important publications that shaped the field
- Help you extract the relevant insights from the literature

<!----------------------------------------------------------------------------->

## Weekly preparation {-}

Before every class meeting (both lectures and practicals) you need to do the 
assigned homework (delineated in the GitBook chapter for that week). This course 
follows a *flipped classroom* procedure, so you must complete the weekly homework 
to meaningfully participate in, and benefit from, the class meetings.

---

### Background knowledge {- #background}

We assume you have basic knowledge about multivariate statistics before entering 
this course. You do not need any prior experience working with R. If you wish to 
refresh your knowledge, we recommend the chapters on ANOVA, multiple regression,
and exploratory factor analysis from Field’s [*Discovering Statistics using R*][field]. 

If you cannot access the Field book, many other introductory statistics textbooks 
cover these topics equally well. So, use whatever you have lying around from 
past statistics courses. You could also try one of the following open-access options:

- [*Applied Statistics with R*][stats_with_r]
- [*Introduction to Modern Statistics*][modern_stats]
- [*Introduction to Statistical Learning*][isl]

<!----------------------------------------------------------------------------->

## Grading {-}

Your grade for the course is based on a “portfolio” composed of the three 
take-home assignments:

1. Latent variable modeling 
    - Deadline: **2022-10-07 23:59**
    - Group assignment
    - Contributes 25% of your course grade

1. Path modeling 
    - Deadline: **2022-10-28 23:59**
    - Group assignment
    - Contributes 25% of your course grade

1. Full Structural equation modeling 
    - Deadline: **2022-11-11 23:59**
    - Individual assignment
    - Contributes 50% of your course grade

**The specifics of the assignments are explicated in the 
[Assignments](assignments.html) chapter of this GitBook**

<!----------------------------------------------------------------------------->

## Attendance {-}

Attendance is not mandatory, but we strongly encourage you to attend all lectures 
and practicals. In our experience, students who actively participate tend to pass 
the course, whereas those who do not participate tend to drop out or fail. The 
lectures and practicals build on each other, so, in the unfortunate event that 
you have to miss a class meeting, please make sure you have caught up with the 
material before the next session.

<!----------------------------------------------------------------------------->

[gs]: https://scholar.google.com/
[uu_lib]: https://www.uu.nl/en/university-library
[r]: https://www.r-project.org/
[lavaan]: https://www.lavaan.ugent.be/
[field]: https://www.discoveringstatistics.com/books/discovering-statistics-using-r/
[stats_with_r]: https://book.stat420.org/
[modern_stats]: https://openintro-ims.netlify.app/
[isl]: https://www.statlearning.com

<!--chapter:end:index.Rmd-->

# Assignments {-}

This chapter contains the details and binding information about the three 
assignments that comprise the portfolio upon which your course grade is based.

- For each assignment, you will use R to analyze some real-world data, and you 
will write up your results in a concise report.
    - The required components of these analyses/reports are delineated in the 
    following three sections.
    - You will submit this report via Blackboard.
- You will complete the first two assignments in your [Assignment Group][group].
- You will complete the third assignment individually.
- The first two assignments each contribute 25% of your course grade.
- The third assignment contributes 50% of your course grade.

<!----------------------------------------------------------------------------->

## Assignment 1: Latent Variable Model {-}

In the first assignment, **you will work in groups** to apply a *latent variable 
model* to a real-world problem wherein the observed variables are indirect 
indicators of an unobserved social scientific construct. The components of the 
first assignment are described below.

1. Find a suitable dataset, and describe the data. **(150 words)**.
    a. Data you have collected yourself (e.g., for a previous course)
    a. Open data (e.g., data provided with a published paper)
    a. The “Coping with COVID-19” dataset (if you can’t find anything else)
1. State the research question, and define the theoretical latent variable 
model. **(150 words)**
    a. Be sure to clearly explicate all parts of your model.
1. Translate your theoretical latent variable model into `lavaan` syntax, and 
estimate the model.
    a. Provide relevant output in a suitable format.
1. Explain your rationale for important modeling decisions. **(300 words)**
    a. Motivate your choice of latent variable model (i.e., EFA, CFA, PCA).
    a. Discuss assumptions.
    a. Discuss other important decisions that could have influence your results.
1. Report and interpret the results in APA style.
    a. Include measures of model fit.
1. Discuss the results. **(300 words)**
    a. Use your results to answer the research question.
    a. Consider the strengths and limitations of your analysis.
     
---

### Submission {-}

Assignment 1 is due at 23:59 on Friday 7 October 2022

- Submit your report via the [Assignment 1][a1] portal on Blackboard.

<!----------------------------------------------------------------------------->

## Assignment 2: Path Model {-}

For the second assignment, **you will work in groups** to apply a *path model* 
that describes how several variables could be causally related. The components 
of the second assignment are described below.

1. Find a suitable dataset, and describe the data. **(150 words)**
    a. Data you have collected yourself (e.g., for a previous course)
    a. Open data (e.g., data provided with a published paper)
    a. The “Coping with COVID-19” dataset (if you can’t find anything else)
1. State the research question, and define the theoretical path model. 
**(150 words)**
    a. This model can be a re-analysis of a question that was originally 
    tested using regression or ANOVA in a published paper.
1. Translate your theoretical path model into `lavaan` syntax, and estimate the 
model.
    a. Provide relevant output in a suitable format.
1. Explain your rationale for important modeling decisions. **(300 words)**
    a. Discuss the conceptual fit between your theory and your model.
    a. Evaluate the model assumptions.
    a. If applicable, discuss the differences and similarities between your
    path model and the regression or ANOVA results from the original paper.
    a. Discuss other important decisions that could have influence your results.
1. Report and interpret the results in APA style.
    a. Include measures of explained variance for the dependent variables.
1. Discuss the results. **(300 words)**
    a. Use your results to answer the research question.
    a. Consider the strengths and limitations of your analysis.

---

### Submission {-}

Assignment 2 is due at 23:59 on Friday 28 October 2022

- Submit your report via the [Assignment 2][a2] portal on Blackboard.

<!----------------------------------------------------------------------------->

## Assignment 3: Full Structural Equation Model {-}

In the third assignment, **you will work individually** to apply a full *SEM* 
that describes how several (latent) variables could be causally related. The 
components of the third assignment are described below.

1. Find a suitable dataset, and describe the data. **(150 words)**
    a. Data you have collected yourself (e.g., for a previous course)
    a. Open data (e.g., data provided with a published paper)
    a. The “Coping with COVID-19” dataset (if you can’t find anything else)
1. State the research question, and define the theoretical SEM. 
**(200 words)**
    a. This model can be a re-analysis of a question that was originally tested 
    using regression or ANOVA in a published paper.
    a. The structural component of this model must include, at least, three
    variables.
    a. The model must include, at least, one latent variable.
1. Translate your theoretical SEM into `lavaan` syntax, and estimate the model.
    a. Provide relevant output in a suitable format.
1. Explain your rationale for important modeling decisions. **(300 words)**
    a. Discuss the conceptual fit between your theory and your model.
    a. Evaluate the model assumptions.
    a. If applicable, discuss the differences and similarities between your path 
    model and the regression or ANOVA results from the original paper.
    a. Discuss other important decisions that could have influence your results.
1. Report and interpret the results in APA style.
    a. Include measures of model fit.
    a. Include measures of explained variance for the dependent variables.
1. Discuss the results. **(500 words)**
    a. Use your results to answer the research question.
    a. Consider the strengths and limitations of your analysis.

---

### Submission {-}

Assignment 3 is due at 23:59 on Friday 11 November 2022

- Submit your report via the [Assignment 3][a3] portal on Blackboard.

<!----------------------------------------------------------------------------->

## Data {-}

If you have your own data, that's great, and we strongly encourage you to use
those data for the projects. That being said, most students will not have their
own data. If you're one of the majority without data, you'll have to find an 
existing dataset to analyze (in case you were wondering, you won't have time to 
collect new data).

---

### Sources {-}

Those who need data can make use of one of the myriad publicly available 
datasets available online. Of course these datasets will vary widely in terms of 
their applicability to the project and the amount of work required to prepare 
the variables for analysis. 

The following links point to various sources of publicly available data.

- [Kaggle datasets](https://www.kaggle.com/datasets)
- [GSS Sata Explorer](https://gssdataexplorer.norc.org/) 
- [StatLine Open Data](https://www.cbs.nl/en-gb/our-services/open-data/)
- [Open Data from the Dutch Government](https://data.overheid.nl/)
- [Google Cloud Platform](https://console.cloud.google.com/marketplace/browse?filter=solution-type:dataset&_ga=2.82807711.826450092.1580463730-1425634986.1574932949)
- [Google dataset search](https://toolbox.google.com/datasetsearch) 

---

### Tips {-}

When searching for a suitable dataset, consider the following tips:

- It will take longer than you expect.
    - Start searching for data early.
    - Plan on devoting a good chunk of time to tracking down a useful dataset.
    - Realize that even after you find the data, you will also have to do a good 
    bit of cleaning to get the variables ready for analysis.
- Expect to be disappointed. 
    - You may find a dataset that looks good on paper but won't work for the 
    project (e.g., too much missing data, poor coverage for key relations, jenky 
    distributions, etc.).
        - You usually won't discover these issue until you've already downloaded 
        the data and started exploring/cleaning.
        - Make sure you check the actual data you've downloaded before you 
        declare success in your search.
- Keep the project requirements in mind when searching for data. 
    - You will need, at least, three interesting variables to model in some sort 
    of causal process (required for Assignment 3).
    - You will need, at least, one latent variable (required for Assignments 1 & 3).
        - To define a latent variable, you need multiple items (at least three) 
        measuring the same hypothetical construct.
        - In practice, these multiple indicators usually come from scales (e.g., 
        NEO-PI Big Five Personality Inventory, Beck Depression Inventory).
- If you find a dataset that looks suitable, but you want to confirm, send 
me ([Kyle](mailto:k.m.lang@uu.nl)) a link to the dataset, and I'll let you know 
if it will work for the assignments.
        
<!----------------------------------------------------------------------------->

## Procedures {-}

---

### Formating {-}

- You must submit your assignment reports in PDF format.
- Each report should include a title page.
- The title page should include the following information:
    - The name of the assignment.
    - The names of all assignment authors (i.e., all group members for 
    Assignments 1 & 2, your name for Assignment 3).
    - The Assignment Group number (only for Assignments 1 & 2).

---

### Length {-}

In the preceding three sections, some line-items in the descriptions of the 
assignments concluded with a word count.

- These sections of your report must conform to the indicated word counts.
    - E.g., the *Data* section of Assignment 1 must contain no more that 150 
    words, and the *Discussion* section of Assignment 3 must contain no more 
    than 500 words.
- In all other sections, you may use as many words as necessary to adequately 
explain yourself (though concision and parsimony are still encouraged).

Note that the assignments are **not** intended to be full-blown papers! You only 
get a few hundred words to describe your data, justify your theoretical model, 
and discuss the results. The focus should be on the definition of your model, 
how this model relates to theory (introduction), and what you have learned from 
your estimated model (discussion).

---

### Submission {-}

You will submit your reports through Blackboard.

- Each assignment has a corresponding item in the "Assignments" section of the 
BB page through which you will submit your reports.
- For Assignments 1 & 2, you may only submit one report per group.
    - Designate one group member to submit the report.
    - The grade for this submission will apply to all group members.
- If something goes wrong with the submission, or you notice a mistake (before 
the deadline) that you want to correct, you may upload a new version of your 
report. 
    - We will grade the final submitted version.
- The submissions will be screened with SafeAssign.

<!----------------------------------------------------------------------------->

## Grading {-}

Each assignment comprises 6 components (i.e., the line-items labelled 1--6 in 
the assignment descriptions).

- Each of these 6 elements will be scored as:
    - Insufficient (0 points)
    - Sufficient (1 point)
    - Excellent (1.5 points)
    
The individual assignment grades will be computed as follows:

1. Sum the 6 component grades.
1. Add 1 to the sum of the component grades.

Consequently, we have the following landmarks:

- All components marked as *Insufficient* $\rightarrow$ Assignment Grade = 1
- All components marked as *Sufficient* $\rightarrow$ Assignment Grade = 7
- All components marked as *Excellent* $\rightarrow$ Assignment Grade = 10

The final course grade, $G_{course}$, will be computed as the weighted average of 
the three individual assignment grades, $G_{a1}$, $G_{a2}$, $G_{a3}$:

\[
G_{course} = 0.25 G_{a1} + 0.25 G_{a2} + 0.5 G_{a3}
\]

<!----------------------------------------------------------------------------->

## Rules {-}

---

### Resources {-}

For all three assignments, you may use any reference materials you like, 
including:

- All course materials
- The course GitBook
- Additional books and papers
- The internet

---

### Collaboration {-}

You will complete the first two assignments in groups.

- Although you will work in groups, your group may not work together with other 
groups.

You will complete the final assignment individually. 

- For this assignment, you may not work with anyone else.

For all three assignments, you are obligated to submit original work (i.e., work 
conducted for this course by you or your group). 

- Submitting an assignment that violates this condition constitutes fraud.
- Such cases of fraud will be addressed according to the 
[University's standard policy](https://students.uu.nl/en/practical-information/policies-and-procedures/fraud-and-plagiarism).

---

### Academic integrity {-}

Hopefully, you also feel a moral obligation to obey the rules. For this course, 
we have implemented an examination that allows you to showcase what you have 
learned in a more realistic way than a written exam would allow. 

- This assessment format spares you the stress of long exams (the two exams for 
this course used to be 4 hours each) and the attendant studying/cramming. 
- The assignments will also help you assess your ability to independently 
analyse data, which is important to know for your future courses and/or career.

However, this format also assumes that you complete the assignments in good 
faith. So, I simply ask that you hold up your end of the bargain, and submit 
your original work to show us what you’ve learned.

---

### Strict stuff {-}

By submitting your assignments (both group and individual), you confirm the 
following:

1. You have completed the assignment yourself (or with your group)
1. You are submitting work that you have written yourself (or with your group)
1. You are using your own UU credentials to submit the assignment
1. You have not had outside help that violates the conditions delineated above 
while completing the assignment

All assignments will be submitted via SafeAssign in Blackboard and, thereby, 
checked for plagiarism. If fraud or plagiarism is detected or suspected, we will 
inform the Board of Examiners in the usual manner. In the event of demonstrable 
fraud, the sanctions delineated in Article 5.15 of the 
[Education and Examination Regulations (EER)](https://students.uu.nl/en/fss/education-and-examination-regulations-eer)
will apply. 

<!----------------------------------------------------------------------------->

[a1]: https://uu.blackboard.com/webapps/assignment/uploadAssignment?content_id=_4261056_1&course_id=_141610_1
[a2]: https://uu.blackboard.com/webapps/assignment/uploadAssignment?content_id=_4261057_1&course_id=_141610_1
[a3]: https://uu.blackboard.com/webapps/assignment/uploadAssignment?content_id=_4261058_1&course_id=_141610_1
[group]: https://uu.blackboard.com/webapps/bb-group-mgmt-LEARN/execute/groupInventoryList?course_id=_141610_1

<!--chapter:end:assignments.Rmd-->

```{r, include = FALSE}
library(dplyr)
library(foreign)

dataDir  <- "../../../data/"
imageDir <- "../images/"
partDir  <- "sections/"
```

<!--
\Sexpr{knit_child("w0_setup.Rmd", quiet = TRUE) %>% cat(sep = "\n")}
\Sepxr{knit_child("w0_basics.Rmd", quiet = TRUE) %>% cat(sep = "\n")}
\Sexpr{knit_child("w0_tutorial.Rmd", quiet = TRUE) %>% cat(sep = "\n")}
-->

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "w0_setup.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w0_tutorial.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
```

<!--chapter:end:week0.Rmd-->

# Introduction

```{r week1_global_options, include = FALSE}
library(dplyr)

dataDir  <- "../../../data/"
quizDir  <- "../assess/"
partDir  <- "sections/"
imageDir <- "../images/"
```

**Homework before the lecture**

1. Complete the preparatory material:
    a. Read over the [Course Information](index.html) chapter
    a. Work through the [Software Setup](software-setup.html) chapter
    a. Work through the [R Tutorial](r-tutorial.html) chapter
1. Watch the [Lecture Recording](lecture.html) for this week.
1. Complete the [Reading](reading.html) for this week, and answer the associated 
reading questions.
1. Complete the [Formative Assessment](formative-assessment.html) for this week.

**Lecture content**

We start with a brief introduction to the course, the course goals and rules, 
and the general idea of statistical modeling. 

- We will introduce the type of models that we will consider in this course.
- We will shortly discuss several related concepts:
    - Model simplicity/complexity
    - Model fit
    - Graphical representations of model parameters
    - Interpretations of model parameters   

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "w1_lecture.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w1_reading.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w1_assess.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w1_home.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w1_class.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
```

<!--chapter:end:week1.Rmd-->

# EFA

```{r week2_global_options, include = FALSE}
library(dplyr)

dataDir  <- "../../../data/"
quizDir  <- "../assess/"
partDir  <- "sections/"
imageDir <- "../images/"
```

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-1.html) for this week.
1. Complete the [Reading](reading-1.html) for this week, and answer the 
associated reading questions.
1. Complete the [Formative Assessment](formative-assessment-1.html) for this week.

**Lecture content**

This lecture constitutes a general introduction to latent variables and scaling 
procedures. We will discuss several different aspects of exploratory factor 
analysis (EFA). Most notably:

- The differences between Principal Component Analyses (PCA) and Factor Analysis
- Model estimation and factor extraction methods
- Factor rotations

You will have to make decisions regarding each of these aspects when conducting
a factor analysis. We will also discuss reliability and factor scores as means 
of evaluating the properties of a scale.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-1.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-1.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "w2_lecture.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_reading.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_assess.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_home.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_class.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
```

<!--chapter:end:week2.Rmd-->

