# Introduction to R

```{r week1_global_options, include = FALSE}
source("code/chapter_setup.R")
partDir <- paste0(partDir0, "week1/")
```

This week, we will cover basic R and RStudio usage.

**Homework before the lecture**

1. Complete the preparatory material:
    a. Read over the [Course Information](index.html) chapter
    a. Work through the [Software Setup](software-setup.html) chapter
    a. Download the [Course Data](course-data.html) from SurfDrive
    <!-- a. Work through the [R Tutorial](r-tutorial.html) chapter -->
1. Watch the [Lecture Recordings](lecture.html) for this week.
<!------------------------------------------------------------------------------ 
1. Complete the [Reading](reading.html) for this week, and answer the associated 
reading questions.
1. Complete the [Formative Assessment](formative-assessment.html) for this week.
------------------------------------------------------------------------------->

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knit_child(paste0(partDir, "lecture.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
# knit_child(paste0(partDir, "../reading-placeholder.Rmd"), quiet = TRUE) %>%
#   cat(sep = "\n")
# knit_child(paste0(partDir, "../assess-placeholder.Rmd"), quiet = TRUE) %>%
#   cat(sep = "\n")
knit_child(paste0(partDir, "home.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
knit_child(paste0(partDir, "class.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
```
