# Full SEM

```{r week6_global_options, include = FALSE}
source("code/chapter_setup.R")
partDir <- paste0(partDir0, "week6/")
```

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-5.html) for this week.
1. Complete the [Reading](reading-5.html) for this week, and answer the 
associated reading questions.
1. Complete the [Formative Assessment](formative-assessment-5.html) for this week.

**Lecture content**

This week, you will focus on integrating all of the disparate methods we've
covered so far into full-fledged *structural equation models*.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-5.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-5.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knit_child(paste0(partDir, "../lecture-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "../reading-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "../assess-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "../home-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "../class-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
```
