# CFA

```{r week5_global_options, include = FALSE}
source("code/chapter_setup.R")
partDir <- paste0(partDir0, "week5/")
```

This week, we will introduce *confirmatory factor analysis* (CFA) and discuss 
how it differs from EFA. Furthermore, we will revisit the ideas of model fit, 
model complexity, and model identification. 

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-4.html) for this week.
1. Complete the [Reading](reading-4.html) for this week, and answer the 
associated reading questions.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-4.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-4.html).
 
**Extra resources**

If you're still struggling with some of the content after working through the
above materials, head over to the [Extra Resources](extra-resources.html) page
and take a look at the embedded tutorial.

```{r, echo = FALSE, message = FALSE, warning = FALSE, error = TRUE, results = "asis"}
knit_child(paste0(partDir, "lecture.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
# knit_child(paste0(partDir0, "lecture-placeholder.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
knit_child(paste0(partDir, "reading.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
knit_child(paste0(partDir, "home.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
knit_child(paste0(partDir, "class.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
knit_child(paste0(partDir, "extra.Rmd"), quiet = TRUE) %>% cat(sep = "\n")
```
