# EFA

```{r week2_global_options, include = FALSE}
library(dplyr)

dataDir  <- "../../../data/"
quizDir  <- "../assess/"
partDir  <- "sections/"
imageDir <- "../images/"
```

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-1.html) for this week.
1. Complete the [Reading](reading-1.html) for this week, and answer the 
associated reading questions.
1. Complete the [Formative Assessment](formative-assessment-1.html) for this week.

**Lecture content**

This lecture constitutes a general introduction to latent variables and scaling 
procedures. We will discuss several different aspects of exploratory factor 
analysis (EFA). Most notably:

- The differences between Principal Component Analyses (PCA) and Factor Analysis
- Model estimation and factor extraction methods
- Factor rotations

You will have to make decisions regarding each of these aspects when conducting
a factor analysis. We will also discuss reliability and factor scores as means 
of evaluating the properties of a scale.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-1.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-1.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "w2_lecture.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_reading.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_assess.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_home.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w2_class.Rmd"), quiet = TRUE) %>% 
  cat(sep = "\n")
```
