# CFA

```{r week5_global_options, include = FALSE}
source("code/chapter_setup.R")
partDir <- paste0(partDir0, "week5/")
```

This week, we will introduce *confirmatory factor analysis* (CFA) and discuss 
how it differs from EFA. Furthermore, we will revisit the idea of model fit and 
introduce into the R-package **lavaan**. 

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-4.html) for this week.
1. Complete the [Reading](reading-4.html) for this week, and answer the 
associated reading questions.
<!-- 1. Complete the [Formative Assessment](formative-assessment-4.html) for this week. -->

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-4.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-4.html).

 
```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knit_child(paste0(partDir, "lecture.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "reading.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
# knit_child(paste0(partDir, "../assess-placeholder.Rmd"), quiet = TRUE) %>%
#   cat(sep = "\n")
knit_child(paste0(partDir, "../home-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knit_child(paste0(partDir, "../class-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
```
