# Mediation & Moderation

```{r week3_global_options, include = FALSE}
library(dplyr)

dataDir  <- "../../../data/"
quizDir  <- "../assess/"
partDir  <- "sections/"
imageDir <- "images/"
codeDir  <- "../code/"
```

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-2.html) for this week.
1. Complete the [Reading](reading-2.html) for this week, and answer the 
associated reading questions.
1. Complete the [Formative Assessment](formative-assessment-2.html) for this week.

**Lecture content**

In this lecture, we will discuss two particular types of processes that we can
model using path analysis: *mediation* and *moderation*.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-2.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-2.html).

<!------------------------------------------------------------------------------
```{r, eval = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "w5_lecture.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w5_reading.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w5_assess.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w5_home.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "w5_class.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
```
------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::knit_child(paste0(partDir, "lecture-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "reading-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "assess-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "home-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
knitr::knit_child(paste0(partDir, "class-placeholder.Rmd"), quiet = TRUE) %>%
  cat(sep = "\n")
```
