## Lecture

One of the many benefits of R is its community of users and developers, and the enormous quantity of learning materials that have been developed for teaching new users. For today's lecture [R Basic Basics](https://rladiessydney.org/courses/ryouwithme/01-basicbasics-0/), we turn to the group R-Ladies Sydney who have produced a series of in-depth online learning resources oriented towards beginners. This series of three lectures covers everything that we need to know for this course to get started using R. 

### Basic Basics 1

```{r, echo = FALSE}
knitr::include_url("https://www.youtube.com/embed/kfcX5DEMAp4")
```

This first lecture above introduces the layout of RStudio, projects and management, setting up and saving scripts, running your code, managing your environment, 

Optionally, use the [website](https://rladiessydney.org/courses/ryouwithme/01-basicbasics-1/) to follow along.

### Basic Basics 2

```{r, echo = FALSE}
knitr::include_url("https://www.youtube.com/embed/v6VygIgvoZU")
```

The second lecture demonstrates installing and loading packages, and how to add comments to a script.

Optionally, use the [website](https://rladiessydney.org/courses/ryouwithme/01-basicbasics-2/) to follow along.


### Basic Basics 3

```{r, echo = FALSE}
knitr::include_url("https://www.youtube.com/embed/2MVolYETR5Q")
```

Finally, in the third lecture, we learn how to load data from a file, and explore it using some basic functions. 

The data set from the lecture [sydneybeaches.csv](https://raw.githubusercontent.com/rladiessydney/RYouWithMe/master/sydneybeaches.csv)

Optionally, use the [website](https://rladiessydney.org/courses/ryouwithme/01-basicbasics-3/) to follow along.
