## Formative Assessment

```{r include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

fName <- paste0(imageDir, "formative2.png")

if(!file.exists(fName)) {
  library(tidySEM)
  library(lavaan)
  library(ggplot2)

  df        <- iris[ , 1:4]
  names(df) <- c("y_1", "x", "y_2", "y_3")
  
  set.seed(58)
  
  tidy_sem(df) |>
    measurement() |>
    estimate_lavaan() |>
    graph_sem(fix_coord = TRUE) -> p

  ggsave(fName, p, "png")
}
```

A formative assessment helps you judge your progress in the course and helps you 
address any blind spots in your understanding of the material. 

- This assessment does not contribute to your course grade.
- If you get a question wrong, you will receive a hint on how to improve your 
understanding of the material.

Ideally, you should complete this assessment  *after* you have watched the 
lecture recording, but *before* the lecture meeting. 

- This way, if you discover gaps in your understanding, we can discuss any 
topics that need more attention during the lecture meeting.

---

```{r results = "asis", echo = FALSE}
source(paste0(quizDir, "formative.R"))
questionnaire(paste0(quizDir, "formative2.csv"))
```
