## In-Class Exercises

```{r include = FALSE}
opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
dataDir <- "../../../../data/"
```

During this practical, you will work through some exercises meant to expand your
statistical reasoning skills and improve your understanding of linear models.

For this exercise, having some familiarity with regression will be helpful. If
you feel like you need to refresh your knowledge in this area, consider the
resources listed in the [Background knowledge](#background) section.

**Data:**

You will use the following dataset for these exercises.

- [Sesam.sav](https://surfdrive.surf.nl/files/index.php/s/93FzuSnGidunEVA/download?path=%2F&files=Sesam.sav)

--------------------------------------------------------------------------------

### Data Exploration

--------------------------------------------------------------------------------

Open the file "Sesam.sav"

```{r, eval = FALSE}
# Load `dplyr` for data processing:
library(dplyr)

# Load the `haven` library for reading in SPSS files:
library(haven)

## Load the 'Sesam.sav' data
## Use haven::zap_formats() to remove SPSS attributes
sesam <- read_sav(file = "Sesam.sav") %>% zap_formats()
```

```{r, echo = FALSE}
library(dplyr)
library(haven)
sesam <- read_sav(paste0(dataDir, "Sesam.sav")) %>% zap_formats()
```

--------------------------------------------------------------------------------

This file is part of a larger dataset that evaluates the impact of the first
year of the Sesame Street television series. Sesame Street is mainly concerned
with teaching preschool related skills to children in the 3--5 year age range.

The following variables will be used in this exercise:

- **age**: measured in months
- **prelet**: knowledge of letters before watching Sesame Street (range 0--58)
- **prenumb**: knowledge of numbers before watching Sesame Street (range
  0--54)
- **prerelat**: knowledge of size/amount/position relationships before
  watching Sesame Street (range 0--17)
- **peabody**: vocabulary maturity before watching Sesame Street (range
  20--120)
- **postnumb**: knowledge of numbers after a year of Sesame Street (range
  0--54)

*Note*: Unless stated otherwise, the following questions refer to the `sesam`
data and the above variables.

--------------------------------------------------------------------------------

#### 

What is the type of each variable?

*Hint:* The output of the `str()` function should be helpful here.

<details>

<summary>Click to show code</summary>

```{r}
## Examine the data structure:
str(sesam)
```

  <details>
  
  <summary>Click for explanation</summary>
  
  All variables are numeric. `str()` uses the abbreviation "num" to indicate a
  numeric vector.
  
  </details>

</details>

--------------------------------------------------------------------------------

####

- What is the average age in the sample?
- What is the age range (youngest and oldest child)?

*Hint:* Use `tidySEM::descriptives()`

<details>

<summary>Click to show code</summary>

As in the take home exercises, you can use the `descriptives()` function from
the `tidySEM` package to describe the data:

```{r}
library(tidySEM)

descriptives(sesam)
```

  <details>
  
  <summary>Click for explanation</summary>
  
  We can get the average age from the "mean" column in the table (
  `r round(mean(sesam$age), 1)`), and the age range from the columns "min" and
  "max", (`r round(min(sesam$age), 1)` and `r round(max(sesam$age), 1)`
  respectively.)
  </details>

</details>

--------------------------------------------------------------------------------

#### {#changeScore}

- What is the average gain in knowledge of numbers?
- What is the standard deviation of this gain?

*Hints:*

- You will need to compute the gain and save the change score as a new object.
- You can then use the base-R functions mean() and sd() to do the
  calculations.

<details>

<summary>Click to show code</summary>

Create a new variable that represents the difference between pre- and post-test
scores on knowledge of numbers:

```{r}
sesam <- mutate(sesam, ndif = postnumb - prenumb)
```

Compute the mean and SD of the change score:

```{r}
sesam %>% summarise(mean(ndif), sd(ndif))
```

</details>

--------------------------------------------------------------------------------

####

- Create an appropriate visualization of the gain scores you computed in 
  \@ref(changeScore).
- Justify your choice of visualization.

*Hint:* Recall the visualizations for continuous variables covered in the
[Week 1 In-Class Exercises](https://kylemlang.com/r-primers/class.html).

<details>

<summary>Click to show code</summary>

```{r}
library(ggplot2)

## Create an empty baseline plot object:
p <- ggplot(sesam, aes(x = ndif))

## Add some appropriate geoms:
p + geom_histogram()
p + geom_density()
p + geom_boxplot()
```
  <details>
  
  <summary>Click for explanation</summary>
  
  Because the gain score is numeric, we should use something appropriate for
  showing the distribution of a continuous variable. In this case, we can use
  either a density plot, or a histogram (remember from the lecture, this is
  like a density plot, but binned).
    
  We can also use a box plot, which can be a concise way to display a lot of
  information about a variable in a little less space.
  
  </details>

</details>

--------------------------------------------------------------------------------

####

- Create a visualization that provides information about the bivariate
  relationship between the pre- and post-test number knowledge.
- Justify your choice of visualization.
- Describe the relationship based on what you see in your visualization.

*Hint:* Again, some of the visualizations covered in the
[Week 1 In-Class Exercises](https://kylemlang.com/r-primers/class.html) may be
useful.

<details>

<summary>Click to show code</summary>

```{r}
## Create a scatterplot of the pre- and post-test number knowledge
ggplot(sesam, aes(x = prenumb, y = postnumb)) +
  geom_point()
```

  <details>
  
  <summary>Click for explanation</summary>
  
  A scatterplot is a good tool for showing patterns in the way that two
  continuous variables relate to each other. From it, we can quickly gather
  information about whether a relationship exists, its direction, its
  strength, how much variation there is, and whether or not a relationship
  might be non-linear.
  
  Based on this scatterplot, we see a positive relationship between the prior
  knowledge of numbers and the knowledge of numbers at the end of the study.
  Children who started with a higher level of numeracy also ended with a 
  higher level of numeracy. 
    
  There is a considerable amount of variance in the relationship. Not every
  child increases their numeracy between pre-test and post-test. Children show
  differing amounts of increase. 
  
  </details>

</details>

--------------------------------------------------------------------------------

### Linear Modeling

--------------------------------------------------------------------------------

#### 

Are there significant, bivariate associations between *postnumb* and the
following variables?

- *age*
- *prelet*
- *prenumb*
- *prerelat*
- *peabody*

Use Pearson correlations to answer this question.

- You do not need to check the assumptions here (though you would in real life).

*Hint:* The base-R `cor.test()` function and the `corr.test()` function from the
`psych` package will both conduct hypothesis tests for a correlation
coefficients (the base-R `cor()` function only computes the coefficients).

<details>

<summary>Click to show code</summary>

```{r}
library(psych)

## Test the correlations using psych::corr.test():
sesam %>%
  select(postnumb, age, prelet, prenumb, prerelat, peabody) %>%
  corr.test()

## OR ##

library(magrittr)

## Test the correlations using multiple cor.test() calls:
sesam %$% cor.test(postnumb, age)
sesam %$% cor.test(postnumb, prelet)
sesam %$% cor.test(postnumb, prenumb)
sesam %$% cor.test(postnumb, prerelat)
sesam %$% cor.test(postnumb, peabody)
```

  <details>
  
  <summary>Click for explanation</summary>
  
  Yes, based on the p-values (remember that 0 here really means very small,
  making it less than .05), we would say that there are significant
  correlations between postnumb and all other variables in the data. (In fact,
  all variables in the data are significantly correlated with one another.)
  
  </details>

</details>

--------------------------------------------------------------------------------

#### {#reg1}

Do *age* and *prenumb* explain a significant proportion of the variance in
*postnumb*?

- What statistic did you use to justify your conclusion?
- Interpret the model fit.
- Use the `lm()` function to fit your model.

<details>

<summary>Click to show code</summary>

```{r}
lmOut <- lm(postnumb ~ age + prenumb, data = sesam)
summary(lmOut)
```

```{r, include = FALSE}
s <- summary(lmOut)

r2 <- s$r.squared %>% round(3)
f <- s$fstatistic
df1 <- f[2]
df2 <- f[3]
f <- f[1] %>% round(3)
```

  <details>

  <summary>Click for explanation</summary>
  - Yes, *age* and *prenumb* explain a significant amount of variability in
    *postnumb* ($R^2 = `r r2`$, $F[`r df1`, `r df2`] = `r f`$, $p < 0.001$).
  - We use the F statistic for the overall test of model fit to support this
    conclusion.
  - The variables *age* and *prenumb* together explain `r 100 * r2`% of the
    variability in *postnumb*.

  </details>

</details>

--------------------------------------------------------------------------------

####

Write the null and alternative hypotheses tested for in \@ref(reg1).

<details>

<summary>Click for explanation</summary>

Since we are testing for explained variance, our hypotheses concern the $R^2$.

$$
\begin{align*}
H_0: R^2 = 0\\
H_1: R^2 > 0
\end{align*}
$$

Note that this is a directional hypotheses because the $R^2$ cannot be negative.

</details>

--------------------------------------------------------------------------------

#### {#syntax1}

Define the model syntax to estimate the model from \@ref(reg1) as a path
analysis using `lavaan`.

<details>

<summary>Click to show code</summary>

```{r}
mod <- 'postnumb ~ 1 + age + prenumb'
```

</details>

--------------------------------------------------------------------------------

#### {#path1}

Estimate the path analytic model you defined above.

- Use the `lavaan::sem()` function to estimate the model.

<details>

<summary>Click to show code</summary>

```{r}
library(lavaan)

lavOut1 <- sem(mod, data = sesam)
```

</details>

--------------------------------------------------------------------------------

####

Summarize the fitted model you estimated above.

- Use the `summary()` function to summarize the model.

<details>

<summary>Click to show code</summary>

```{r}
summary(lavOut1)
```

</details>

--------------------------------------------------------------------------------

In OLS regression, the predictor variables are usually treated as fixed and do
not covary. We can easily relax this assumption in path analysis.

--------------------------------------------------------------------------------

####

Re-estimate the path analytic model you defined in \@ref(syntax1).

- Specify the predictors as random, correlated variables.
  
*Hint:* You can make the predictors random in, at least, two ways:
    
1. Modify the model syntax to specify the correlation between `age` and 
   `prenumb`.
1. Add `fixed.x = FALSE` to your `sem()` call.

<details>

<summary>Click to show code</summary>

```{r}
lavOut2 <- sem(mod, data = sesam, fixed.x = FALSE)

## OR ##

mod <- '
postnumb ~ 1 + age + prenumb
age ~~ prenumb
'

lavOut2 <- sem(mod, data = sesam)
```

</details>

--------------------------------------------------------------------------------

####

Summarize the fitted model you estimated above.

- Compare the results to those from the OLS regression in \@ref(reg1) and the
  path model in \@ref(path1).

<details>

<summary>Click to show code</summary>

```{r}
summary(lavOut2)
summary(lavOut1)
summary(lmOut)
```

</details>

--------------------------------------------------------------------------------

####

Consider the path model below.

- How many regression coefficients are estimated in this model?
- How many variances are estimated?
- How many covariances are estimated?

```{r, echo = FALSE}
mod <- "
postnumb ~ prerelat + prelet + prenumb
prerelat ~ age
prelet ~ age
prenumb ~ age
"

res <- lavaan::sem(mod, data = sesam)

prepare_graph(res,
  layout = get_layout("",    "prerelat", "",
                      "age", "prelet",   "postnumb",
                      "",    "prenumb",  "",
    rows = 3
  ),
  angle = 1,
  curvature = 60) %>% 
  if_edges({show == TRUE}, {label = NA}) %>% 
  if_edges({to == "postnumb" & op == "~"}, {connect_to = c("top", "left", "bottom")}) %>%
  if_edges({to == "postnumb" & op == "~~"}, {connect_to = "right"; connect_from = "right"}) %>% 
  plot()

```

<details>
  
<summary>Click for explanation</summary>

- Six regression coefficients (red)
- Four (residual) variances (blue)
- No covariances

```{r, echo = FALSE}
  mod <- "
  postnumb ~ prerelat + prelet + prenumb
  prerelat ~ age
  prelet ~ age
  prenumb ~ age
  "
  
res <- lavaan::sem(mod, data = sesam)
prepare_graph(res,
              layout = get_layout("",    "prerelat", "",
                                  "age", "prelet",   "postnumb",
                                  "",    "prenumb",  "",
                                  rows = 3
              ),
              angle = 1,
              curvature = 60) %>% 
  if_edges({show == TRUE}, {label = NA}) %>% 
  if_edges({to == "postnumb" & op == "~"}, {connect_to = c("top", "left", "bottom")}) %>%
  if_edges({to == "postnumb" & op == "~~"}, {connect_to = "right"; connect_from = "right"}) %>% 
  color_reg("red") %>%
  color_var("blue") %>%
  edit_nodes({
    color <- "black"
  }) %>%
  #if_nodes(condition = {
  #  name == "age"
  #}, expr = {
  #  color <- "orange"
  #}) %>% 
  plot(p)
```

</details>

--------------------------------------------------------------------------------

####

Consider a multiple regression analysis with three continuous independent
variables: scores on tests of language, history, and logic, and one continuous
dependent variable: score on a math test. We want to know if scores on the
language, history, and logic tests can predict the math test score.

- Sketch a path model that you could use to answer this question
- How many regression parameters are there?
- How many variances could you estimate?
- How many covariances could you estimate?

--------------------------------------------------------------------------------

### Categorical IVs

--------------------------------------------------------------------------------

Load the `Drivers.sav` data.

```{r, eval = FALSE}
# Read the data into a data frame named 'drivers':
drivers <- read_sav("Drivers.sav") %>%
  as_factor() # This preserves the SPSS labels for nominal variables
```

```{r, echo = FALSE}
drivers <- read_sav(paste0(dataDir, "Drivers.sav")) %>%
  as_factor()
```

--------------------------------------------------------------------------------

In this section, we will evaluate the following research question:

    Does talking on the phone interfere with people's driving skills?

These data come from an experiment. The *condition* variable represents the
three experimental conditions:

- Hand-held phone
- Hands-free phone
- Control (no phone)

We will use *condition* as the IV in our models. The DV, *RT*, represents the
participant's reaction time (in milliseconds) during a driving simulation.

--------------------------------------------------------------------------------

####

Use the package `ggplot2` to create a density plot for the variable *RT*. 

-   What concept are we representing with this plot? 

*Hint:* Consider the lap times example from 
[the statistical modeling section](#statistical-modeling) of Lecture 2.

<details>

<summary>Click to show code</summary>

```{r}
ggplot(drivers, aes(x = RT)) +
  geom_density()
```
 
  <details>
 
  <summary>Click for explanation</summary>
 
  This shows the distribution of all the combined reaction times from drivers
  in all three categories. 
 
  </details>

</details>

--------------------------------------------------------------------------------

####

Modify this density plot by mapping the variable *condition* from your data to
the `fill` aesthetic in `ggplot`.

- What is the difference between this plot and the previous plot? 
- Do you think there is evidence for differences between the groups?
- How might we test this by fitting a model to our sample?

<details>

<summary>Click to show code</summary>

*Hint*: To modify the transparency of the densities, use the aesthetic `alpha`.

```{r}
ggplot(drivers, aes(x = RT, fill = condition)) +
  geom_density(alpha = .5)
```
 
  <details>
 
  <summary>Click for explanation</summary>
 
  This figure models the conditional distribution of reaction time, where the
  type of cell phone usage is the grouping factor. 
 
  Things you can look at to visually assess whether the three groups differ are
  the amount of overlap of the distributions, how much distance there is between
  the individual means, and whether the combined distribution is much different
  than the conditional distributions.

  If we are willing to assume that these conditional distributions are normally
  distributed and have equivalent variances, we could use a linear model with 
  dummy-coded predictors. 
 
  </details>
 
</details>

--------------------------------------------------------------------------------

#### Aside: ANOVA vs. Linear Regression {.unnumbered}

As you may know, the mathematical model underlying ANOVA is just a linear
regression model with nominal IVs. So, in terms of the underlying statistical
models, there is no difference between ANOVA and regression; the differences lie
in the focus of the analysis.

- ANOVA is really a type of statistical test wherein we are testing hypotheses
  about the effects of some set of nominal grouping factors on some continuous
  outcome.

    - When doing an ANOVA, we usually don't interact directly with the
      parameter estimates from the underlying model.

- Regression is a type of statistical model (i.e., a way to represent a
  univariate distribution with a conditional mean and fixed variance).

    - When we do a regression analysis, we primarily focus on the estimated
      parameters of the underling linear model.

When doing ANOVA in R, we estimate the model exactly as we would for linear
regression; we simply summarize the results differently.

- If you want to summarize your model in terms of the *sums of squares* table
  you usually see when running an ANOVA, you can supply your fitted `lm` object
  to the `anova()` function.

This is a statistical modeling course, not a statistical testing course, so we 
will not consider ANOVA any further.

--------------------------------------------------------------------------------

#### {#reg2}

Estimate a linear model that will answer the research question stated in the 
beginning of this section.

- Use `lm()` to estimate the model.
- Summarize the fitted model and use the results to answer the research question.

<details>

<summary>Click to show code</summary>

```{r}
library(magrittr)

lmOut <- drivers %>%
  mutate(condition = relevel(condition, ref = "control")) %$%
  lm(RT ~ condition)

summary(lmOut)
anova(lmOut)
```

  <details>
  
  <summary>Click for explanation</summary>
  
```{r, include = FALSE}
  tmp <- anova(lmOut)
  
  f <- tmp[1, 4] %>% round(2)
  df1 <- tmp[1, 1]
  df2 <- tmp[2, 1]
  p <- tmp[1, 5] %>% round(3)
```
  
  The effect of *condition* on *RT* is nonsignificant
  ($F[`r df1`, `r df2`] = `r f`$, $p = `r p`$). Therefore, based on these
  results, we do not have evidence for an effect of mobile phone usage on
  driving performance.
  </details>

</details>

--------------------------------------------------------------------------------

####

Use **lavaan** to estimate the model from \@ref(reg2) as a path model.

*Hint*: **lavaan** won't let us use factors for our categorical predictors. So,
you will need to create your own dummy codes.

<details>

<summary>Click to show code</summary>

```{r}
## Specify the model syntax:
mod <- 'RT ~ 1 + HH + HF'

## Dummy code the factor variable:
drivers2 <- mutate(drivers,
                   ## Create dummy code for "hand-held" condition:
                   HH = ifelse(condition == "hand-held", 1, 0),
                   ## Create dummy code for "hands-free" condition:
                   HF = ifelse(condition == "hands-free", 1, 0)
                   )

## Estimate the path model:
lavOut <- sem(mod, data = drivers2)

## Summarize the results:
summary(lavOut)
```

Alternatively, we can combine the data processing and model estimation into a
single pipeline, as shown below. 

- This type of combined approach might be preferable if we do not want to
  permanently add the dummy codes to the `drivers` dataset.
- The multi-step approach shown above would be better if we were going to reuse
  the dummy codes in subsequent analyses.

```{r}
lavOut2 <- drivers %>%
  mutate(HH = ifelse(condition == "hand-held", 1, 0),
         HF = ifelse(condition == "hands-free", 1, 0)
         ) %>%
  sem(mod, data = .) # Estimate the model

## Do the two approaches produce the same results?
all.equal(
  parameterEstimates(lavOut), 
  parameterEstimates(lavOut2)
  )

all.equal(
  fitMeasures(lavOut), 
  fitMeasures(lavOut2)
  )
```

At this point, we haven't covered the tools you need to conduct the ANOVA-style 
tests with path models. So, you can't yet answer the research question with the
above model.

- When we discuss model comparisons, you'll get the missing tools.

</details>

--------------------------------------------------------------------------------

End of In-Class Exercises

--------------------------------------------------------------------------------
