## Formative Assessment

```{r include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

fName <- paste0(imageDir, "formative2.png")

if(!file.exists(fName)) {
  library(tidySEM)
  library(lavaan)
  library(ggplot2)

  df <- cbind(iris[1:75, 1:3], iris[76:150, 2:4])
  lo <- get_layout("x1", 
                   "x2", 
                   "x3", 
                   "x4", 
                   "x5", 
                   "x6",
                   "",   
                   "f1", 
                   "",   
                   "",   
                   "f2", 
                   "", 
                   rows = 2)
  
  tmp1 <- cbind(
    unname(
      expand.grid(c("f1", "f2"), paste0("x", 1:6), stringsAsFactors = FALSE)
      ), 
    arrow = "last")
  tmp2 <- data.frame(cbind(c("f1", "f2"), c("f1", "f2")), arrow = "last")
  tmp3 <- data.frame(cbind(paste0("x", 1:6), paste0("x", 1:6)), arrow = "both")
  
  names(tmp1) <- names(tmp2) <- names(tmp3) <- c("from", "to", "arrow")
  dat         <- rbind(tmp1, tmp2, tmp3)
  
  set.seed(58)
  p <- graph_sem(layout = lo, edges = dat, angle = 179, fix_coord = TRUE)
  
  print(p)
  
  ggsave(fName, p, "png")
}
```

A formative test helps you assess your progress in the course, and helps you address any blind spots in your understanding of the material.
If you get a question wrong, you will receive a hint on how to improve your understanding of the material.

Complete the formative test ideally *after* you've seen the lecture, but *before* the lecture meeting in which we can discuss any topics that need more attention.

```{r results = "asis", echo = FALSE}
source(paste0(quizDir, "formative.R"))
questionnaire(paste0(quizDir, "formative2.csv"))
```

