## In-Class Exercises

In these exercises, we will continue with our re-analysis/replication of the 
Kestilä (2006) results. Rather than attempting a direct replication, we will now
redo the analysis using *exploratory factor analysis* (EFA).

---

### Load the *ESSround1-a.csv* dataset

``` {r, eval = FALSE}
library(foreign)
ess <- read.csv("ESSround1-b.csv", to.data.frame = TRUE)
```

``` {r, echo = FALSE}
#dataDir <- "../../data/"
ess <- read.csv(paste0(dataDir, "ESSround1-b.csv"))

summary(ess)
```

---

###

Kestilä (2006) claimed that running a PCA is a good way to test if the questions 
in the ESS measure attitudes towards immigration and trust in politics.

- Based on what you've learned from the readings and lectures, do you agree with 
this position?

<details>
  <summary>Click for information</summary>

Hopefully not. PCA is not a method for estimating latent measurement structure; 
PCA is a dimension reduction technique that tries to summarize a set of data 
with a smaller set of component scores.

If we really want to estimate the factor structure underlying a set of observed 
variables, we should use EFA.

</details>

---

###

Suppose you had to construct the *trust in politics* and *attitude towards 
immigration* scales described by Kestilä (2006) based on the theory and 
background information presented in that article.

- What type of analysis would you choose? 
- What key factors would influence your decision?

<details>
  <summary>Click for information</summary>
  
- We are trying to estimate meaningful latent factors, so EFA would be an 
appropriate method.
- The theory presented by Kestilä (2006) did not hypothesize a particular number
of factors, so we would need to use appropriate techniques to estimate the best
number. In particular, combining information from:
    - Scree plots
    - Parallel analysis
    - Substantive interpretability of the (rotated) factor loadings
- Since the factors are almost certainly correlated, we should apply an oblique 
rotation.

</details>

---

We will now rerun the two PCAs that you conducted for the 
[At-Home Exercises](at-home-exercises-1.html) using EFA.

We need to know how many factors to extract to estimate an EFA using the 
`psych::fa()` function.

- We could simply estimate a range of solutions and compare the results.
- We can restrict the range of plausible solutions and save some time by first 
checking/plotting the eigenvalues and running parallel analysis.

---

###

Estimate the number of latent factors underlying the *Trust* items based on the 
eigenvalues, the scree plot, and parallel analysis.

- How many factors are suggested by each method?

*Hints*:

- You can estimate the eigenvalues of a data matrix with the `eigen()` function 
from Base-R.
    - Check the *Value* section of the documentation for `eigen()` to find 
    information on how to extract the eigenvalues.
- You can create a scree plot by supplying the vector of eigenvalues to the
`qplot()` function from the `ggplot2` package and applying the `geom_path()` geometry.

<details>
  <summary>Click for explanation</summary>

First, we'll run a trivial EFA using the `psych::fa()` function to get eigenvalues.

- We don't care about the factors yet, so we can extract a single factor.
- We also don't care about interpretable solutions, so we don't need rotation.

```{r}
## Load the psych package:
library(psych)

## Run a trival EFA on the 'trust' items
efa_trust0 <- fa(ess[7:19], nfactors = 1, rotate = "none")
```

We can check the eigenvalues to see what proportion of the observed variance is 
accounted for by each additional factor we may extract.

```{r}
round(efa_trust0$values, digits = 3)
```

Since only one eigenvalue is greater than one, the so-called "Kaiser Criterion" 
would suggest extracting a single factor. 

- The Kaiser Criterion is not a valid way to select the number of factors in EFA. 
So, we don't want to rely on this information alone.

We can still use the eigenvalues to help us with factor enumeration, though. One
way to do so is by plotting the eigenvalues in a *scree plot*. 

- We can easily create a scree plot by passing the eigenvalues to the `qplot()` 
function from the `ggplot2` package and add apply the `geom_path()` geometry:

```{r}
library(ggplot2)

qplot(y = efa_trust0$values) + 
  geom_path() +
  ylab("Eigenvalues") + 
  xlab("Factors")
```

Although the scree plot provides useful information, we need to interpret that 
information subjectively, and the conclusions are sometimes ambiguous. 

- In this case, the plot seems to suggest either one or three components, 
depending on where we consider the "elbow" to lie.

As recommended in the lecture, we can also use "parallel analysis" (Horn, 1965) 
to provide more quantified information about the number of factors. We'll use 
the `psych::fa.parallel()` function to implement parallel analysis.

- Parallel analysis relies on randomly simulated/permuted data, so we should set 
a seed to make sure our results are reproducible.
- We can set the `fa = "fa"` option to get only the results for EFA.

```{r}
## Set the random number seed:
set.seed(235711)

## Run the parallel analysis:
pa_trust <- fa.parallel(ess[7:19], fa = "fa")
```

The results of the parallel analysis suggest `r pa_trust$nfact` factors. 

---

If you've been paying close attention, you may have noticed that we need to
compute the eigenvalues from the original data to run parallel analysis. Hence,
we don't actually need to run a separate EFA to estimate the eigenvalues.

```{r}
## View the eigenvalues estimated during the parallel analysis:
pa_trust$fa.values

## Compare to the version from the EFA:
pa_trust$fa.values - efa_trust0$values

## Recreate the scree plot from above:
qplot(y = pa_trust$fa.values) + 
  geom_path() + 
  ylab("Eigenvalues") + 
  xlab("Factors")
```

---

The different criteria disagree on how many factors we should extract, but we
have narrowed the range.

- Based on the scree plot and parallel analysis, we should consider solutions 
for 3 to `r pa_trust$nfact` factors. 
- We need to examine the factor loadings to see which solution makes the most 
substantive sense.

</details>

---

###

Do the same analysis for the *attitudes toward immigration* items.

<details>
  <summary>Click for explanation</summary>

This time, we'll start by running the parallel analysis and get the eigenvalues 
from the object returned by `psych::fa.parallel()`.

```{r}
## Set the seed:
set.seed(235711)

## Run parallel analysis on the 'attitudes' items:
pa_att <- fa.parallel(ess[20:44])

## Check the eigenvalues:
round(pa_att$fa.values, digits = 3)

## Create a scree plot:
qplot(y = pa_att$fa.values) + 
  geom_path() +
  ylab("Eigenvalues") +
  xlab("Factors")
```

For the *attitudes toward immigration* analysis, the results are even more 
ambiguous than they were for the *trust* items.

- The Kaiser Criterion suggests `r sum(pa_att$fa.value > 1)` factors
- The scree plot is hopelessly ambiguous
    - At least 3 factors?
    - No more than 9 factors?
- Parallel analysis suggests `r pa_att$nfact` factors

Based on the scree plot and parallel analysis, it seems reasonable to consider 
solutions for 3 to `r pa_att$nfact` factors.

- Again, we need to check the substantive interpretation to choose the most 
reasonable solution.

</details>

---

To evaluate the substantive interpretability of the different solutions, we need
to estimate the full EFA models for each candidate number of factors. We then 
compare the factor loadings across solutions to see which set of loading define
the most reasonable set of latent variables.

---

### {#trust_efa}

For the *trust* items, estimate the EFA models for each plausible number of 
components that you identified above.

- Use the `psych::fa()` function to estimate the models.

You will need to specify a few key options.

- The data (including only the variables you want to analyze)
- The number of factors that you want to extract
- The rotation method
- The estimation method
- The method of estimating factor scores

*Hint:* You can save yourself a lot of typing/copy-pasting (and the attendant 
chances of errors) by using a `for()` loop to iterate through numbers of factors.

<details>
  <summary>Click for explanation</summary>

```{r}
## Define an empty list to hold all of our fitted EFA objects:
efa_trust <- list()

## Loop through the interesting numbers of factors and estimate an EFA for each:
for(i in 3:6) 
  efa_trust[[i - 2]] <- fa(ess[7:19], 
                           nfactors = i,          # Number of factors = Loop index
                           rotate   = "promax",   # Oblique rotation
                           scores   = "Bartlett") # Estimate factor scores with WLS
```

</details>

---

### {#att_efa}

Repeat the analysis from \@ref(trust_efa) for the *attitudes* items.

<details>
  <summary>Click for explanation</summary>
 
```{r}
efa_att <- list()
for(i in 3:7)
  efa_att[[i - 2]] <- fa(ess[20:44], 
                         nfactors = i,         
                         rotate   = "promax",  
                         scores   = "Bartlett")
```

</details>

---

###

Compare the factor loading matrices from the models estimated in \@ref(trust_efa), 
and select the best solution for the *Trust* model.

*Hints:*

- The factor loadings are stored in the `loadings` slot of the object returned 
by `psych::fa()`.
- Looping can also be useful here.

<details>
  <summary>Click for explanation</summary>
 
```{r}
for(x in efa_trust) print(x$loadings)
```

Note that any factor loadings with magnitude lower than 0.1 are suppressed in
above output.

The factor loadings matrix indicates how strongly each latent factor (columns) 
associates with the observed items (rows). We can interpret these factor loadings 
in the same way that we would interpret regression coefficients (indeed, a factor 
analytic model can be viewed as a multivariate regression model wherein the 
latent factors are the predictors and the observed items are the outcomes). 

- A higher factor loading indicates a stronger association between the item and 
factor linked by that loading. 
    - Items with high factor loadings are "good" indicators of the respective 
    factors.
- Items with only very low loadings do not provide much information about any factor.
    - You may want to exclude such items from your analysis.
    - Note that the size of the factor loadings depends on the number of factors. 
    So, you should consider excluding an observed item after you have chosen the
    number of latent factors.

When we print the loading matrix, we see additional information printed below
the factor loadings.

- *Proportion Var*: What proportion of the items' variance is explained by each 
of the factors.
- *Cumulative Var*: How much variance the factors explain, in total. 
    - If you estimated as many factors as items, then the *Cumulative Var* for
    the final factor would be `1.00` (i.e., 100%).

</details>

---

###

Compare the factor loading matrices from the models estimated in \@ref(att_efa), 
and select the best solution for the *Attitudes* model.

<details>
  <summary>Click for explanation</summary>
 
```{r}
for(x in efa_att) print(x$loadings)
```

</details>

---

It is very possible that you selected a different numbers of factors than Kestilä 
(2006). We need to keep these exercises consistent, though. So, the remaining 
questions will all assume you have extract three factors from the *Trust* items
and five factors from the *Attitudes* items, to parallel the Kestilä (2006) results.

```{r}
## Select the three-factor solution for 'trust':
efa_trust <- efa_trust[[1]]

## Select the five-factor solution for 'attitudes':
efa_att <- efa_att[[3]]
```

---

### {#factorScores}

Give the factor scores meaningful names, and add the scores to the `ess` dataset 
as new columns.

*Hint:* If you're not sure of what do to, check \@ref(pcScores).

<details>
  <summary>Click for explanation</summary>

```{r}
## Rename the factor scores:
colnames(efa_trust$scores) <- c("trust_pol", "satisfy", "trust_inst")
colnames(efa_att$scores)   <- c("effects", 
                                "allowance", 
                                "refugees", 
                                "ethnic", 
                                "europe")

## Add factor scores to the dataset as new columns:
ess <- data.frame(ess, efa_trust$scores, efa_att$scores)
```

</details>

---

Kestilä (2006) used the component scores to descriptively evaluate country-level 
differences in *Attitudes toward Immigration* and *Political Trust*. So, now it's 
time to replicate those analyses.

---

###

Repeat the Kestilä (2006) between-country comparison using the factor scores you 
created in \@ref(factorScores) and an appropriate statistical test.

<details>
  <summary>Click for explanation</summary>

Here, we'll only demonstrate a possible approach to analyzing one of the *Trust*
dimensions. We can use an ANOVA to test whether the countries differ in average 
levels of *Trust in Institutions* (as quantified by the relevant factor score).

```{r}
out <- aov(trust_inst ~ cntry, data = ess)
summary(out)
```

According to the omnibus F-test, average levels of *Trust in Institutions* 
significantly differ between countries, but this test cannot tell us between 
which countries the differences lie. 

One way to test for differences between the individual countries would be a 
posthoc test of all pairwise comparisons. Since we'll be doing `r choose(10, 2)` 
tests, we need to apply a correction for repeated testing. Below, we use Tukey's 
HSD correction.

```{r}
TukeyHSD(out)
```

</details>

---

The second part of the Kestilä (2006) analysis was to evaluate how 
socio-demographic characteristics affected attitudes towards immigrants and 
trust in politics among the Finlish electorate. Before we can replicate this part
of the analysis, we need to subset the data to only the Finnish cases.

---

###

Create a new data frame that contains only the Finnish cases from `ess`.

*Hint:* You can use logical indexing based on the `cntry` variable. 

<details>
  <summary>Click for explanation</summary>
  
```{r}
ess_finland <- ess[ess$cntry == "Finland", ]
```

</details>

---

We still have one more step before we can estimate any models. We must prepare 
our variables for analysis.

- Our dependent variables will be the factor scores generated above. So, we do 
not need to apply any further processing.
- We have not yet used any of the indepdendent variables, though. So, we should 
inspect those variables to see if they require any processing. 

In the `ess` data, the relevant variables have the following names:

- `gndr`
- `yrbrn`
- `eduyrs`
- `polintr`
- `lrscale`

---

###

Inspect the independent variables listed above.

<details>
  <summary>Click for explanation</summary>

```{r}
library(tidySEM)
descriptives(ess_finland[, c("gndr", "yrbrn", "eduyrs", "polintr", "lrscale")])
```

It looks like we still need some recoding. 

**1. Age**

The data contain the participants' years of birth instead of their age, but 
Kesilä analyzed age. Fortunately, we know that the data were collected in 2002, 
so we can simply subtract each participant's value of `yrbrn` from the 2002 to
compute their age.


```{r}
ess_finland$age <- 2002 - ess_finland$yrbrn
```

**2. Political Interest & Political Orientation**

The variables `polintr` and `lrscale` are represented as character vectors. If
we analyze these variables as they are, R will convert the character vectors to 
factors and make dummy codes for each distinct value; we definitely don't want that. 

We need to convert `polintr` and `lrscale` to numeric vectors, but we cannot
naively apply the `as.numeric()` function, because R doesn't know how to convert
a word into a number (or, at least, how we want the operation to be done). There
are many ways that we could go about this conversion, but the `recode()` function 
from the `dplyr` package is particularly useful arbitrary recoding tasks such as 
we have here.

Kestilä (2006) dichotomized `polintr` by combining the lowest two and highest 
two categories. So, we don't actually want to convert the `polint` variable into 
a numeric, Likert-type variable. We want `polint` to be a binary factor. The 
`recode_factor()` function from `dplyr()` will automatically convert our result
into a factor.

```{r}
## Load the dplyr package:
library(dplyr)

## Recode the four character values into two factor levels:
ess_finland$polintr_bin <- recode_factor(ess_finland$polintr,
                                         "Not at all interested" = "Low Interest",
                                         "Hardly interested" = "Low Interest",
                                         "Quite interested" = "High Interest",
                                         "Very interested" = "High Interest")

## Check the conversion:
table(new = ess_finland$polint_bin, old = ess_finland$polintr)
```

For political orientation, only the two extreme values of the scale are labeled 
as text. So, we only need to recode those two levels.

- Since we're only replacing two of the 11 values, we'll need to provide a
value for the `.default` argument in `dplyr::recode()`.

```{r}
## Recode the extreme levels:
tmp <- recode(ess_finland$lrscale,
              "Left" = 0,
              "Right" = 10,
              .default = as.numeric(ess_finland$lrscale)
              )

## Check the conversion:
table(new = tmp, old = ess_finland$lrscale)

## Overwrite the old variable:
ess_finland$lrscale <- tmp
```

**3. Sex**

Although `gndr` would be automatically converted to a factor by `lm()`, it's 
better to be explicit about our intentions and create the factor ourselves.

```{r}
ess_finland$sex <- factor(ess_finland$gndr)
```

</details>

---

Now, we're finally ready to replicate the regression analysis from Kestilä (2006).
Creating a single aggregate score by summing the individual component scores is 
a pretty silly thing to do, though. So, we won't reproduce that aspect of the 
analysis.

---

###

Run a series of multiple linear regression analyses with the factor scores you 
created in \@ref(factorScores) as the dependent variables and the same predictors 
used by Kestilä (2006). 

- Do your results agree with those reported by Kestilä (2006)?

```{r}
## Predicting 'Trust in Institutions':
out_trust_inst <- lm(trust_inst ~ sex + age + eduyrs + polintr_bin + lrscale, 
                     data = ess_finland)
summary(out_trust_inst)

## Predicting 'Trust in Politicians':
out_trust_pol <- lm(trust_pol ~ sex + age + eduyrs + polintr_bin + lrscale, 
                    data = ess_finland)
summary(out_trust_pol)

## Predicting 'Attitudes toward Refugees':
out_refugees <- lm(refugees ~ sex + age + eduyrs + polintr_bin + lrscale, 
                   data = ess_finland)
summary(out_refugees)
```

</details>

---

That does it for our replication of the Kesilä (2006) analyses, but we still have
one more topic to consider in this practical. One of the most common applications
of EFA is *scale development*. Given a pool of items without a known factor
structure, we try to estimate the underlying latent factors that define the
(sub)scales defined by our items.

In such applications, we use the factor loading matrix for our optimal solution 
to make "bright-line" assignments of items to putative factors according to the 
simple structure represented by the estimated factor loading matrix. In other 
words, we disregard small factor loadings and assign observed items to only the
single latent factor upon which they load most strongly. We then hypothesize that
those items are true indicators of that latent factor.

We can use *confirmatory factor analysis* (which you will learn about next week)
to test rigorously this hypothesis, but we can already get started by estimating
the *internal consistency* (a type of reliability) of the hypothesized subscales. 

---

###

Estimate the internal consistency of the three *Trust* subscales and five 
*Attitudes* subscales implied by your EFA solutions from \@ref(efa_trust) and \@ref(efa_att).

- Use *Cronbach's Alpha* to quantify internal consistency.
- Use the `alpha()` function from the `psych()` package to conduct the analysis.

The next step is to assert whether the items that together form one factor, also form a reliable scale. Run separate reliability analyses by means of the function `alpha` for the items that together form one factor, and evaluate Cronbach's alpha to see whether the scales are internally consistent. The "Reliability if an item is dropped (alpha.drop)" information may be handy to inspect what would happen if you would delete one item; you can find it inside the reliability analysis object. If Cronbach's alpha is not ok, deselect one survey item and run the analyses under question 4 and question 5 again.

*Hint: Cronbach's alpha > .7 are deemed to be ok, > .8 is good.* 

If Cronbach's alpha is not ok, deselect one survey item and run the analyses under question 4 and question 5 again.

<details>
  <summary>Click for explanation</summary>

If you want to assess the reliability of the variables `pltcare`, `pltinvt`, `trstprl`, `trstplt`, and `trstep` you can run a reliability analysis as follows. 

*Hint: name the new objects substantively, instead of numbering them.*

```{r}
reli_1 <- psych::alpha(ess[, c(7, 8, 9, 12, 13)])
reli_1
```

Sometimes, the table "Reliability if an item is dropped" will indicate that Cronbach's alpha increases when you drop a variable out of the analysis. Note, however, that doing so is an exploratory approach to analysis, and it may make your work incomparable to other publications using the same scale.

`r if(knitr::is_html_output()){"\\details"}`



###

Save your syntax and your data, you will need it next week.
