## In-Class Exercises

```{r echo = FALSE, warning=FALSE, message=FALSE}
library(lavaan)
library(semTools)
library(dplyr)
library(magrittr)
```

As in the At-Home Exercises, these exercises are based on the following paper.

[
Metha, A., Chen, E, Mulvenon, S., and Dode, I. (1998). A theoretical model of 
suicide risk. *Archives of Suicide Research, 4*, 115--133.
](https://doi.org/10.1080/13811119808260442)

We will now pick-up where we left off with the At-Home Exercises.

---

### 

Load the *suicide_risk.rds* dataset.

<details>
  <summary>Click for explanation</summary>

```{r, echo = FALSE}
suicide <- readRDS(paste0(dataDir, "suicide_risk.rds"))
```

```{r, eval = FALSE}
suicide <- readRDS("suicide_risk.rds")
```

</details>

---

According to Metha et al (1998), the process of suicide risk is more complicated 
than what we modeling in the At-Home Exercises. On page 117--118 the authors 
summarize their expectations in five hypotheses. These hypotheses imply a path
model that describes how the four continuous predictors relate to *suicide risk*. 

---

### {#ieSyntax1}

Write the **lavaan** model syntax for the path model implied by the hypotheses
of Metha et al (1998). 

- Leave *sex* out of the model for now. 

<details>
  <summary>Click for explanation</summary>

```{r}
mod <- '
suirisk ~ depression + b * hopeless + subabuse
hopeless ~ a * depression + selfesteem
selfesteem ~ depression

ie := a*b
'
```

</details>

---

###

Use `lavaan::sem()` to estimate the model.

- Does the model fit well?
- Are the hypothesized effects significant?

<details>
  <summary>Click for explanation</summary>

```{r fit1, cache = TRUE}
library(lavaan)

## First fit without bootstrapping to check model adequacy:
fit <- sem(mod, data = suicide)
summary(fit, fit.measures = TRUE)

## Now bootstrap to get appropriate test of indirect effect:
set.seed(235711)
fit_boot <- sem(mod, data = suicide, se = "boot", bootstrap = 1000)
```

```{r, echo = FALSE}
tmp <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
```

The model fits the data well 
($\chi^2[`r tmp[2]`] = `r tmp[1] %>% round(2)`$, 
$p = `r tmp[3] %>% round(3)`$,
$\textit{CFI} = `r tmp[4] %>% round(3)`$,
$\textit{TLI} = `r tmp[5] %>% round(3)`$,
$\textit{RMSEA} = `r tmp[6] %>% round(3)`$,
$\textit{SRMR} = `r tmp[7] %>% round(3)`$), and all but the second hypothesis 
are supported.

```{r, echo = FALSE}
tmp <- summary(fit)$pe %>% filter(op == "~")

h1 <- tmp %>% filter(lhs == "suirisk", rhs == "depression")
h1a <- tmp %>% filter(label == "a")
h1b <- tmp %>% filter(label == "b")

h3 <- tmp %>% filter(lhs == "selfesteem", rhs == "depression")
h4 <- tmp %>% filter(lhs == "hopeless", rhs == "selfesteem")
h5 <- tmp %>% filter(lhs == "suirisk", rhs == "subabuse")

ie <- summary(fit_boot, ci = TRUE)$pe %>% filter(label == "ie")
```

- H1: The direct effect of *depression* on *suicide risk* is significant 
($\beta = `r h1$est %>% round(3)`$,
$z = `r h1$z %>% round(2)`$,
$p < 0.001$).
- H2: The indirect effect of *depression* on *suicide risk* through 
*hopelessness* is not significant 
($\hat{\textit{IE}} = `r ie$est %>% round(3)`$,
$95\%~CI = [`r ie$ci.lower %>% round(3)`; ~ `r ie$ci.upper %>% round(3)`]$).
    - *Depression* significantly predicts *hopelessness*, after controlling for 
    *self-esteem* 
    ($\beta = `r h1a$est %>% round(3)`$,
    $z = `r h1a$z %>% round(2)`$,
    $p < 0.001$).
    - *Hopelessness* does not significantly predict *suicide risk*, after 
    controlling for *depression* and *substance abuse* 
    ($\beta = `r h1b$est %>% round(3)`$,
    $z = `r h1b$z %>% round(2)`$,
    $p = `r h1b$p %>% round(3)`$).
- H3: *Depression* significantly predicts decreases in *self-esteem* 
($\beta = `r h3$est %>% round(3)`$,
$z = `r h3$z %>% round(2)`$,
$p < 0.001$).
- H4: *Self-esteem* significantly predicts *hopelessness* 
($\beta = `r h4$est %>% round(3)`$,
$z = `r h4$z %>% round(2)`$,
$p < 0.001$).
- H5: *Substance abuse* significantly predicts *suicide risk*, after controlling
for *depression* and *hopelessness* 
($\beta = `r h5$est %>% round(3)`$,
$z = `r h5$z %>% round(2)`$,
$p < 0.001$).

</details>

---

After failing to support H2, Metha et al (1998) modified there model based on 
post hoc analyses. They made two changes to the model structure:

1. Added a path from *hopelessness* to *substance abuse*.
1. Removed the path from *hopelessness* to *suicide risk*.

---

### 

Modify the model syntax from \@ref(ieSyntax1) to reflect the post hoc 
modifications described above.

<details>
  <summary>Click for explanation</summary>

```{r}
mod <- '
suirisk ~ depression + subabuse
hopeless ~ depression + selfesteem
subabuse ~ hopeless
selfesteem ~ depression
'
```

</details>

---

### {#modifiedModel}

Use `lavaan::sem()` to estimate the modified model.

- Does the model fit well?
- Are the structural paths significant?

<details>
  <summary>Click for explanation</summary>

```{r}
## First fit without bootstrapping to check model adequacy:
fit <- sem(mod, data = suicide)
summary(fit, fit.measures = TRUE)
```

```{r, echo = FALSE}
tmp <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
```

Yes, the modified model also fits the data well 
($\chi^2[`r tmp[2]`] = `r tmp[1] %>% round(2)`$, 
$p = `r tmp[3] %>% round(3)`$,
$\textit{CFI} = `r tmp[4] %>% round(3)`$,
$\textit{TLI} = `r tmp[5] %>% round(3)`$,
$\textit{RMSEA} = `r tmp[6] %>% round(3)`$,
$\textit{SRMR} = `r tmp[7] %>% round(3)`$), and all structural paths are now 
statistically significant.

</details>

---

Although the modified model seems to be a good representation of the data, the
original and modified models imply substantially different theoretical processes.
Think about what the two models say regarding the process by which *depression* 
influences *suicide risk* and the role that *hopelessness* and *substance abuse*
play therein. We will discuss these ideas in the plenary lecture session.

---

Sex-based differences were a key component of the Metha et al (1998) analysis.
In the original paper, *sex* was conceptualized as a moderator, and the
differences induced thereby were evaluated by analyzing the male and female
subsamples separately.

As a general rule, subsample analysis is rarely advisable (especially in SEM). 
We hypothesize moderation by *sex*, but the subsample analysis only allows us to
infer this moderation indirectly through qualitative comparisons of the 
sex-specific estimates. We will be much better off directly modeling the influence
of *sex* by specifying a multiple-group model.

In **lavaan**, we can easily estimate any model as an unrestricted multiple-group
model by adding the `group = "grouping_variable_name"` to the fitting function.
You can find a tutorial at <https://lavaan.ugent.be/tutorial/groups.html>.

---

### {#fullMgModel}

Re-estimate the model from \@ref(modifiedModel) as a multiple-group model with 
`sex` as the grouping factor.

- Specify the same model structure in each group.
- Do not place any between-group constraints on the estimates.
- Does the model fit well?

<details>
  <summary>Click for explanation</summary>

```{r}
fit <- sem(mod, data = suicide, group = "sex")
summary(fit, fit.measures = TRUE)
```

```{r, echo = FALSE}
tmp <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
```

Yes, the model fits the data well 
($\chi^2[`r tmp[2]`] = `r tmp[1] %>% round(2)`$, 
$p = `r tmp[3] %>% round(3)`$,
$\textit{CFI} = `r tmp[4] %>% round(3)`$,
$\textit{TLI} = `r tmp[5] %>% round(3)`$,
$\textit{RMSEA} = `r tmp[6] %>% round(3)`$,
$\textit{SRMR} = `r tmp[7] %>% round(3)`$)

</details>

---

###

Qualitatively compare the effects in each group. Do you notice any salient 
differences?

*Hint:* The standardized estimates are probably useful here. 

<details>
  <summary>Click for explanation</summary>

```{r}
summary(fit, standardized = TRUE)
```

```{r, echo = FALSE}
tmp <- summary(fit, standardized = TRUE)$pe %>% 
  filter(op == "~", rhs == "depression", lhs %in% c("suirisk", "selfesteem"))

b1m <- tmp %>% filter(lhs == "suirisk", group == 1) %>% select("std.all") %>% as.numeric()
b1f <- tmp %>% filter(lhs == "suirisk", group == 2) %>% select("std.all") %>% as.numeric()

b2m <- tmp %>% filter(lhs == "selfesteem", group == 1) %>% select("std.all") %>% as.numeric()
b2f <- tmp %>% filter(lhs == "selfesteem", group == 2) %>% select("std.all") %>% as.numeric()
```

The direct effect of *depression* on *suicide risk* and the effect of 
*depression* on *self esteem* both seem to be quite a bit stronger for males 
($\beta_{SR} = `r b1m %>% round(3)`$, $\beta_{SE} = `r b2m %>% round(3)`$) 
than females
($\beta_{SR} = `r b1f %>% round(3)`$, $\beta_{SE} = `r b2f %>% round(3)`$).

</details>

---

###

Qualitatively compare the $R^2$ for *suicide risk* in the two groups. What do 
you conclude?

<details>
  <summary>Click for explanation</summary>

```{r}
## Extract just the R-Squared estimates with the inspect() function:
inspect(fit, "r2")
```

```{r, echo = FALSE}
tmp <- inspect(fit, "r2")
```

The $R^2$ for males (`r tmp$male["suirisk"] %>% round(3)`) is larger than the 
$R^2$ for females (`r tmp$female["suirisk"] %>% round(3)`). Hence, the model 
better explains suicide risk for males than it does for females.

</details>

---

To test the moderating influence of the grouping variable, we need to test for
significant differences between the group-specific estimates of the structural 
paths. If a path differs significantly between the two groups, we conclude that 
the grouping factor *moderates* that path.

We can conceptualize these test in two general ways.

1. An omnibus test: Does group affect any of the structural paths?
1. Path-specific tests: Does group affect an individual path?

To implement an omnibus test for moderation by group, we simply need to compare 
the fit of the full, unrestricted model (such as the one we estimated in 
\@ref(fullMgModel)) to the fit of a restricted model wherein each structural 
path is constrained to be equal in both groups.

The `group.equal` argument in **lavaan** fitting functions makes this process 
very easy.

---

###

Conduct an omnibus test to see if *sex* moderates any of the paths in the model 
from \@ref(fullMgModel).

- Does *sex* have a significant, overall moderating influence on the model?

*Hints:*

- Check the documentation for `sem()` to see how you need to specify the 
`group.equal` argument.
- You can use the `anova()` function to compare the full and restricted models.

<details>
  <summary>Click for explanation</summary>

First, estimate the restricted model.

```{r}
fit_res <- sem(mod, 
               data = suicide,
               group = "sex",
               group.equal = "regressions")

## Summarize the results to check that everything went well:
summary(fit_res, fit.measures = TRUE)
```

OK, that looks good. The regression paths are constrained to be equal across 
groups. Now, we need to compare the full and restricted models to see if we've
lost a significant amount of model fit through the constraints.

```{r}
anova(fit, fit_res)
```

```{r, echo = FALSE}
tmp <- anova(fit, fit_res)[2, ]

chi <- tmp["Chisq diff"] %>% as.numeric()
df <- tmp["Df diff"] %>% as.numeric()
```
  
  
The $\Delta \chi^2$ test tells us that we have lost a significant level of fit by 
constraining the paths to equality across groups 
($\Delta \chi^2[`r df`] = `r chi %>% round(2)`$, $p < 0.001$). 
Hence, we can infer overall moderation by *sex*.

</details>

---

Finding significant omnibus moderation is interesting, but we probably want to 
know which paths, in particular, show significant sex-based differences. Hence,
we need to test for path-specific moderation.

We have several options when it comes to implementing such tests. One obvious
idea would be to specify a restricted model that constrains only one path to 
equality and conduct the same type of $\Delta \chi^2$ test that we used for the 
omnibus test. While this approach will certainly work, we would have to specify
and estimate a separate restricted model for every path that we wanted to test.
Thankfully, we have a couple simpler alternatives.

The most direct approach entails the following steps:

1. Label the relevant parameters
1. Define new parameters that represent the differences between the 
group-specific versions of whatever parameters we want to test
1. Estimate the model and evaluate the path-specific moderation by the 
significance tests for these difference parameters

---

### 

Define the **lavaan** model syntax that uses the approach described above to 
test if *sex* moderates the following effects

1. *Depression* $\rightarrow$ *Suicide Risk*
1. *Substance Abuse* $\rightarrow$ *Suicide Risk*
1. *Hopelessness* $\rightarrow$ *Substance Abuse*

<details>
  <summary>Click for explanation</summary>

```{r}
mod <- '
suirisk ~ c(m1, f1) * depression + c(m2, f2) * subabuse
hopeless ~ depression + selfesteem
subabuse ~ c(m3, f3) * hopeless
selfesteem ~ depression

d1 := m1 - f1
d2 := m2 - f2
d3 := m3 - f3
'
```

</details>

---

### {#modTest1}

Estimate the model defined above.

- Does *sex* moderate these three focal effects?

<details>
  <summary>Click for explanation</summary>

```{r}
fit <- sem(mod, data = suicide, group = "sex")
summary(fit)
```

```{r, echo = FALSE}
tmp <- summary(fit)$pe %>% filter(label %in% paste0("d", 1:3))

d1 <- tmp[1, ]
d2 <- tmp[2, ]
d3 <- tmp[3, ]
```

- *Sex* significantly moderates the effect of *depression* on *suicide risk* 
($\beta = `r d1$est %>% round(3)`$,
$z = `r d1$z %>% round(2)`$,
$p = `r d1$p %>% round(3)`$) 
such that males have significantly stronger associations.
- *Sex* also moderates the effect of *hopelessness* on *substance abuse* 
($\beta = `r d3$est %>% round(3)`$,
$z = `r d3$z %>% round(2)`$,
$p = `r d3$p %>% round(3)`$) 
such that males have significantly weaker associations.
- *Sex* does not moderate the effect of *substance abuse* on *suicide risk* 
($\beta = `r d2$est %>% round(3)`$,
$z = `r d2$z %>% round(2)`$,
$p = `r d2$p %>% round(3)`$).

</details>

---

We can also conduct analogous tests without explicitly defining the parameter 
differences as new parameters. To do so, we use the `lavaan::lavTestWald()` 
function and the following procedure.

1. Label the relevant paths
1. Estimate the model
1. Submit the fitted model object and a syntax segment defining the desired 
constraint(s) to `lavTestWald()`.

This approach is nice when we have already estimated the model with the paths 
labeled but no difference parameters defined. In such situations, we can do our
tests without rewriting the model syntax and re-estimating the model.

---

###

Use `lavTestWald()` to test for moderation in the same three paths that you
tested in \@ref(modTest1).

- What are your conclusions?
- Are the results the same as what you get when defining difference parameters?

*Hints:* 

- You do not need to re-specify your model syntax; you can use the fitted model 
from \@ref(modTest1) as input to `lavTestWald()`.
- To get tests of moderation for individual paths, you need to run `lavTestWald()` 
with a single parameter constraint specified.
    - You will need to call `lavTestWald()` three times to replicate the results 
    from \@ref(modTest1).
    
<details>
  <summary>Click for explanation</summary>

```{r}
## Depression -> Suicide Risk:
lavTestWald(fit, constraints = 'm1 == f1')

## Substance Abuse -> Suicide Risk:
lavTestWald(fit, constraints = 'm2 == f2')

## Hopelessness -> Substance Abuse:
lavTestWald(fit, constraints = 'm3 == f3')
```

This approach produces the same inference as the defined parameters approach. In
fact the p-values are identical across the two approaches, and the test 
statistics produced by `lavTestWald()` are equal to the square of those produced 
by the defined parameters approach.

- This latter equivalence arises because the test statistic in the summarized 
`sem()` output is a Z statistic while the test statistic in the `lavTestWald()` 
output is a $\chi^2$ statistic.

</details>

---

Notice that although Metha et al (1998) only hypothesize a single indirect effect,
their model implies several potential indirect effects. Further notice that the
original model and the modified model imply different sets of indirect effects.

---

### {#ieList}

List the potential indirect effects implied by the original and modified models 
from Metha et al (1998).

<details>
  <summary>Click for explanation</summary>

***Original:***

1. *Depression* $\rightarrow$ *Self-Esteem* $\rightarrow$ *Hopelessness* $\rightarrow$ *Suicide Risk*
1. *Depression* $\rightarrow$ *Hopelessness* $\rightarrow$ *Suicide Risk*
1. *Self-Esteem* $\rightarrow$ *Hopelessness* $\rightarrow$ *Suicide Risk*

***Modified:***

1. *Depression* $\rightarrow$ *Self-Esteem* $\rightarrow$ *Hopelessness* $\rightarrow$ *Substance Abuse* $\rightarrow$ *Suicide Risk*
1. *Depression* $\rightarrow$ *Hopelessness* $\rightarrow$ *Substance Abuse* $\rightarrow$ *Suicide Risk*
1. *Self-Esteem* $\rightarrow$ *Hopelessness* $\rightarrow$ *Substance Abuse* $\rightarrow$ *Suicide Risk*

</details>

---

An indirect effect is defined by the product of some set of regression paths.
These underlying regression paths can be moderated by a grouping factor, so 
indirect effects can also be moderated by grouping factors. Such a process is
called *moderated mediation* because some moderator variable influences the 
indirect effects. In our case, we want to know if *sex* influences the indirect 
effects listed above such that males and females show different indirect effects.

We can use the same logic that we applied to test the moderating influence of 
*sex* on individual regression paths to for moderation of the indirect effects.

There is one caveat to this generalization, though. The validity of the Wald test
implemented by the `lavTestWald()` function depends on the tested parameters 
having normal sampling distributions. The sampling distributions for indirect 
effects are not normal, so we should not use the `lavTestWald()` approach to 
evaluate moderation of the indirect effects. Rather, we should use the defined
parameter approach in combination with bootstrapping.

---

### {#ieFreeSyntax}

Modify the syntax for the modified model (i.e., the model estimated in 
\@ref(fullMgModel)) to specify sex-specific versions of the the potential 
indirect effects you listed in \@ref(ieList).

<details>
  <summary>Click for explanation</summary>

```{r}
mod <- '
suirisk ~ depression + c(sr_saM, sr_saF) * subabuse
hopeless ~ c(h_dM, h_dF) * depression + c(h_seM, h_seF) * selfesteem
subabuse ~ c(sa_hM, sa_hF) * hopeless
selfesteem ~ c(se_dM, se_dF) * depression

## Indirect Effects:
ie1_m := se_dM * h_seM * sa_hM * sr_saM
ie2_m := h_dM * sa_hM * sr_saM
ie3_m := h_seM * sa_hM * sr_saM

ie1_f := se_dF * h_seF * sa_hF * sr_saF
ie2_f := h_dF * sa_hF * sr_saF
ie3_f := h_seF * sa_hF * sr_saF
'
```

</details>

---

###

Estimate the model defined in \@ref(ieFreeSyntax).

- Use bootstrapping to evaluate the indirect effects.
- Are the indirect effects significant?

<details>
  <summary>Click for explanation</summary>

```{r ie_free_boot, cache = TRUE}
set.seed(235711)
out <- sem(mod, data = suicide, group = "sex", se = "boot", bootstrap = 1000)
summary(out, ci = TRUE)
```

None of the 95% bootstrapped CIs for the indirect effects contain zero, so all 
three indirect effects are significant in both groups.

</details>

---

### {#ieTestSyntax}

Modify the model syntax from \@ref(ieFreeSyntax) to include the difference 
parameters necessary for testing the moderating influence of *sex* on the indirect
effects.

<details>
  <summary>Click for explanation</summary>

Rather than re-typing the previous syntax, I'll use the `paste()` function to
add on only the new lines.

```{r}
## Add the new parameter definitions to the existing syntax:
mod <- paste(mod, 
'
ie_diff1 := ie1_m - ie1_f
ie_diff2 := ie2_m - ie2_f
ie_diff3 := ie3_m - ie3_f
',
sep = '\n')

## Check the results:
cat(mod)
```

</details>

---

###

Estimate the model defined in \@ref(ieTestSyntax).

- Use bootstrapping to evaluate moderation of the indirect effects.
- Does *sex* moderate the indirect effects?

<details>
  <summary>Click for explanation</summary>

```{r ie_test_boot, cache = TRUE}
set.seed(235711)
out <- sem(mod, data = suicide, group = "sex", se = "boot", bootstrap = 1000)
summary(out, ci = TRUE)
```

All three 95% bootstrapped CIs for the difference parameters contain zero, so 
*sex* does not moderate any of the indirect effects.

</details>

---

End of In-Class Exercises

---

