## Full SEM

In these exercises, you will use full structural equation modeling (SEM) to
evaluate the *Theory of Reasoned Action* (TORA), which is a popular psychological 
theory of social behavior developed by Ajzen and Fishbein. The theory states that 
actual behavior is predicted by behavioral intention, which is in turn predicted 
by the attitude toward the behavior and subjective norms about the behavior. 
Later, a third determinant was added, *perceived behavioral control*. The extent
to which people feel that they have control over their behavior also influences 
their behavior.

---

The data we will use for this practical are available in the *toradata.csv* file. 
These data were synthesized according to the results of Reinecke (1998)'s
investigation of condom use by young people between 16 and 24 years old. 

The data contain the following variables:

- `respnr`: Numeric participant ID
- `behavior`: The dependent variable *condom use* 
    - Measured on a 5-point frequency scale (How often do you...)
- `intent`: A single item assessing behavioral intention 
    - Measured on a similar 5-point scale (In general, do you intend to...). 
- `attit_1`:`attit_3`: Three indicators of attitudes about condom use
    - Measured on a 5-point rating scale (e.g., using a condom is awkward)
- `norm_1`:`norm_3`: Three indicators of social norms about condom use
    - Measured on a 5-point rating scale (e.g., I think most of my friends would 
    use...)
- `control_1`:`control_3`: Three indicators of percieved behavioral control 
    - Measured on a 5-point rating scale (e.g., I know well how to use a condom)
- `sex`: Binary factor indicating biological sex

---

###

Load the data contained in the *toradata.csv* file.
<details>
  <summary>Click for explanation</summary>
  
```{r, eval = FALSE}
condom <- read.csv("toradata.csv", stringsAsFactors = TRUE)
```

```{r, echo = FALSE}
dataDir <- "../../data/"
condom <- read.csv(paste0(dataDir, "toradata.csv"), stringsAsFactors = TRUE)
```

</details>

---

###

The data contain 3 observed indicators of *attitudes*, *norms*, and *control*.
Run a CFA for these three latent variables.

- Correlate the latent factors.
- Do the data support the measurement model for these latent factors?
- Are the three latent factor significantly correlated?
- Is it reasonable to proceed with our evaluation of the TORA theory?

<details>
  <summary>Click for explanation</summary>

```{r}
library(lavaan)

mod_cfa <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
'

fit <- cfa(mod_cfa, data = condom)

summary(fit, fit.measures = TRUE)
```

- Yes, the model fits the data well, so the data support this measurement structure.
- Yes, all three latent variables are significantly, positively correlated.
- Yes. 
   - The measurement structure is supported, so we can use the latent variables
   to represent the respective constructs in our subsequent SEM. 
   - The TORA doesn't actually say anything about the associations between these 
   three factors, but it makes sense that they would be positively associated.
   So, we should find this result comforting.
   
</details>

---

###

Estimate the basic TORA model as an SEM.

- Predict *intention* from *attitudes* and *norms*.
- Predict *condom use* from *intention*. 
- Use the latent versions of *attitudes* and *norms*. 
- Covary the *attitudes* and *norms* factors.
- Does the model fit well?
- Do the estimates align with the TORA?
- How much variance in *intention* and *condom use* are explained by the model?

<details>
  <summary>Click for explanation</summary>

```{r}
mod <- '
## Define the latent variables:
attitudes =~ attit_1 + attit_2 + attit_3
norms     =~ norm_1  + norm_2  + norm_3

## Define the structural model:         
intent   ~ attitudes + norms
behavior ~ intent
'

fit <- sem(mod, data = condom)

summary(fit, fit.measures = TRUE, rsquare = TRUE)
```

```{r, echo = FALSE}
tmp <- inspect(fit, "r2")
```

- Yes, the model still fits the data very well.
- Yes, the estimates all align with the TORA. Specifically, *attitudes* and 
*norms* both significantly predict *intention*, and *intention* significantly 
predicts *condom use*.
- The model explains `r (100 * tmp$intent) %>% round(3)`% of the variance in 
*intention* and `r (100 * tmp$behavior) %>% round(3)`% of the variance in 
*condom use*.

</details>

---

### {#updatedModel}

Update your model to represent the extended TORA model that includes 
*perceived behavioral control*.

- Regress *condom use* onto *perceived behavioral control*.
- Use the latent variable representation of *control*.
- Covary all three exogenous latent factors.
- Does the model fit well?
- Do the estimates align with the updated TORA?
- How much variance in *intention* and *condom use* are explained by the model?

<details>
  <summary>Click for explanation</summary>

```{r}
mod_tora <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
             
intent   ~ attitudes + norms
behavior ~ intent + control
'

fit_tora <- sem(mod_tora, data = condom)

summary(fit_tora, fit.measures = TRUE, rsquare = TRUE)
```

```{r, echo = FALSE}
tmp <- inspect(fit, "r2")
```

- Yes, the model still fits the data very well.
- Yes, the estimates all align with the updated TORA. Specifically, *attitudes* 
and *norms* both significantly predict *intention*, while *intention* and 
*control* both significantly predict *condom use*.
- The model explains `r (100 * tmp$intent) %>% round(3)`% of the variance in 
*intention* and `r (100 * tmp$behavior) %>% round(3)`% of the variance in 
*condom use*.

</details>

---

The TORA model explicitly forbids direct paths from *attitudes* and *norms* to
*behaviors*; these effects should be fully mediated by the *behavioral intention*. 
The theory does not specify how *perceived behavioral control* should affect 
*behaviors*. There may be a direct effect of *control* on *behavior*, or the
effect may be (partially) mediated by *intention*.

---

###

Evaluate the hypothesized indirect effects of *attitudes* and *norms*.

- Include *attitudes*, *norms*, and *control* in your model as in \@ref(updatedModel).
- Does *intention* significantly mediate the effects of *attitudes* and *norms* 
on *behavior*?
    - Don't forget to follow all the steps we covered for testing mediation.
- Are both of the above effects completely mediated?
- Do these results comport with the TORA? Why or why not?

<details>
  <summary>Click for explanation</summary>

```{r boot1, cache = TRUE}
mod <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
             
intent   ~ a1 * attitudes + a2 * norms
behavior ~ b * intent + control + attitudes + norms

ie_att  := a1 * b
ie_norm := a2 * b
'

fit <- sem(mod, data = condom, se = "bootstrap", bootstrap = 1000)

summary(fit, ci = TRUE)
```

- Yes, both indirect effects are significant according to the 95% bootstrapped CIs.
- Yes, both effects are completely moderated by *behavioral intention*. We can
infer as much because the direct effects of *attitudes* and *norms* on *condom 
use* are both nonsignificant.
- Yes, these results comport with the TORA. Both effects are fully mediated, as
the theory stipulates.

</details>

---

In addition to evaluating the significance of the indirect and direct effects, 
we can also take a model-comparison perspective. We can use model comparisons to
test if removing the direct effects of *attitudes* and *norms* on *condom use* 
significantly decreases model fit. In other words, are those paths needed to 
accurately represent the data, or are they "dead weight". 

---

###

Use a $\Delta \chi^2$ test to evaluate the necessity of modeling the direct
effects of *attitudes* and *norms* on *condom use*.

- What is your conclusion?

<details>
  <summary>Click for explanation</summary>

```{r}
anova(fit, fit_tora)
```

The $\Delta \chi^2$ test is not significant. So, we have not lost a significant
amount of fit by fixing the direct effects to zero. In other words, the complete
mediation model explains the data just as well as the partial mediatio model. So,
we should probably prefer the more parsimonious model.

</details>

---

###

Use some statistical means of evaluating the most plausible way to include 
*perceived behavioral control* into the model.

- Choose between the following three options:

1. *control* predicts *behavior* via a direct, un-mediated effect.
1. *control* predicts *behavior* via an indirect effect that is completely 
mediated by *intention*.
1. *control* predicts *behavior* via both an indirect effect through *intention*
and a residual direct effect.

*Hint:* There is more than one way to approach this problem.

<details>
  <summary>Click for explanation</summary>

We can tackle this problem in a few different ways. One possibility entails 
testing the indirect, direct, and total effects.

```{r boot2, cache = TRUE}
## Allow for partial mediation:
mod1 <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
             
intent   ~ attitudes + norms + a * control
behavior ~ b * intent + c * control

ie    := a * b
total := ie + c
'

set.seed(235711)

fit1 <- sem(mod1, data = condom, se = "bootstrap", bootstrap = 1000)
summary(fit1, ci = TRUE)
```

From the above results, we can see that the direct and total effects are both
significant, but the indirect effect is not. Hence, it probably makes the most
sense to include *control* via a direct (non-mediated) effect on *behavior*.

---

We can also approach this problem from a model-comparison perspective. We can
fit models that encode each pattern of constraints and check which one best 
represents the data. 

```{r}
## Force complete mediation:
mod2 <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
             
intent   ~ attitudes + norms + control
behavior ~ intent
'

## Force no mediation:
mod3 <- '
attitudes =~ attit_1   + attit_2   + attit_3
norms     =~ norm_1    + norm_2    + norm_3
control   =~ control_1 + control_2 + control_3
             
intent   ~ attitudes + norms
behavior ~ intent + control
'

## Re-estimate the partial mediation model without bootstrapping:
fit1 <- sem(mod1, data = condom)

## Estimate the two restricted models:
fit2 <- sem(mod2, data = condom)
fit3 <- sem(mod3, data = condom)

## Check the results:
summary(fit1)
summary(fit2)
summary(fit3)

## Do either of the restricted models fit worse than the partial mediation model?
anova(fit1, fit2)
anova(fit1, fit3)
```

The above $\Delta \chi^2$ tests tell us that the full mediation model fits 
significantly worse than the partial mediation model. Hence, forcing full 
mediation by fixing the direct effect to zero is an unreasonable restraint. The 
total effect model, on the other hand, does not fit significantly worse than the 
partial mediation model. So, we can conclude that removing the indirect effect
and modeling the influence of *control* on *behavior* as an un-mediated direct
association represents the data just as well as a model that allows for both 
indirect and direct effects. Hence, we should prefer the more parsimonious total
effects model.

---

While the two tests above lead us to prefer the non-mediated model, we cannot 
directly say that the the complete mediation model fits significantly worse than
the non-mediated model. We have not directly compared those two models, and we 
cannot do so with the $\Delta \chi^2$. We cannot do such a test because these two 
models are not nested: we must both add and remove a path to get from one model 
specification to the other. Also, both models have the same degrees of freedom,
so we cannot define a sampling distribution against which we would compare the 
$\Delta \chi^2$, anyway.

---

We are not completely without options, though. We can use *information criteria*
to compare non-nested models. The two most popular information criteria are the
*Akaike's Information Criterion* (AIC) and the *Bayesian Information Criterion*
(BIC).

```{r}
## Which model is the most parsimonious representation of the data?
AIC(fit1, fit2, fit3)
BIC(fit1, fit2, fit3)
```

When comparing models based on information criteria, a lower value indicates a 
better model in the sense of a better balance of fit and parsimony. The above 
results show that both the AIC and the BIC agree that the no-mediation model is
the best.

---

So, in the end, regardless of how we approach the question, all of our results 
suggest modeling *perceived behavioral control* as a direct, non-mediated
predictor of *condom* use.

</details>

---

Measurement invariance comes in three (useful) levels:

1. *Configural:* The same pattern of free and fixed effects in all groups
1. *Weak:* Configural + Equal factor loadings in all groups
1. *Strong:* Weak + Equal item intercepts in all groups

You can read more about measurement invariance [here][van_de_schoot_et_al] and
[here][putnick_bornstein], and you can find a brief discussion of how to 
conduct measurement invarince testing in **lavaan** [here][mi_tutorial].

---

###

Test for measurement invariance across *sex* groups.

- Test configural, weak, and strong invariance.
- Test for invariance in all three latent factors simultaneously.
- Is full measurement invariance (i.e., up to and including strong invariance)
supported?

<details>
  <summary>Click for explanation</summary>

```{r}
## Estimate the models:
config <- cfa(mod_cfa, 
              data = condom, 
              group = "sex")
weak <- cfa(mod_cfa, 
            data = condom, 
            group = "sex",
            group.equal = "loadings")
strong <- cfa(mod_cfa, 
              data = condom, 
              group = "sex",
              group.equal = c("loadings", "intercepts")
              )

## Check that everything went well:
summary(config)
summary(weak)
summary(strong)

## Test measurement invariance:
anova(config, weak, strong)

## Make sure the strongly invariant model still fits well in an absolute sense:
fitMeasures(strong)
```

Yes, we have been able to establish full measurement invariance. 

- The configurally invariant model fits the data well.
- The $\Delta \chi^2$ tests support the tenability of the weak and strong 
invariance constraints.
- The strongly invariance model still fits the data well.

</details>

---

###

Using a strongly invariant model, conduct an omnibus test to see if *sex* 
moderates any of the regression paths the TORA model.

- What do you conclude.

<details>
  <summary>Click for explanation</summary>
  
```{r}
fit_full <- sem(mod_tora, 
                data = condom, 
                group = "sex", 
                group.equal = c("loadings", "intercepts")
                )

fit_res <- sem(mod_tora, 
               data = condom, 
               group = "sex", 
               group.equal = c("loadings", "intercepts", "regressions")
               )

anova(fit_full, fit_res)
```

Equating all regression paths across groups produces a significant loss of fit.
Therefore, *sex* must moderate at least some of these paths.

</details>

---

End of At-Home Exercises

---

[mi_tutorial]: https://www.lavaan.ugent.be/tutorial/groups.html
[van_de_schoot_et_al]: https://doi.org/10.1080/17405629.2012.686740
[putnick_bornstein]: https://doi.org/10.1016/j.dr.2016.06.004