---
title: "Group 42"
author: ["Zaphod Beeblebrox", "Tricia McMillan", "Marvin T.P. Andy"]

documentclass: article
pdf-engine: pdflatex
papersize: a4
geometry: margin = 20mm
fontfamily: sans
format:
  pdf:
    number-sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lavaan)
library(knitr)
```

\hrule

\vspace{6pt}

For Assignment 2, you will conduct a CFA and report the results via this
worksheet.

- Provide the requested information for each of the following items.
- Only include the necessary information; waffling will hurt your grade.
- Replace the placeholder values above with your Assignment 2 Group's details
  and the appropriate date.
- Submit the completed worksheet as a PDF document.

*NOTE:* The relative weights for each item are given in brackets.

\vspace{12pt}

\hrule

# Which dataset are you using? [1]

# Describe your hypothesized measurement model. [2]

- *What latent factors are you estimating?*
- *Which items will you use to define each latent factor?*
- *Why is your hypothesized factor structure justified?*

# Provide a path diagram of your theoretical measurement model. [2]

# Provide the lavaan syntax for your model. [2]

```{r}

```

# Use the `lavaan::cfa()` function to estimate your model. [2]

```{r}

```

# Report the relevant parameter estimates from your fitted model. [3]

- *Don't include the entire model summary.*
- *The `lavaan::lavInspect()` or `lavaan::parameterEstimates()` functions could
  be useful here.*

```{r, echo = FALSE}

```

# Report the $R^2$ estimates from your fitted model. [2]

```{r, echo = FALSE}

```

# Report suitable model fit indices for your estimated model. [3]

- *Don't include every fit index lavaan provides.*

```{r, echo = FALSE}

```

# Do your results support the validity of your hypothesized model? [4]

- *Explain why or why not.*

# Do you think the model could be improved? [2]

- *If so, explain how you would modify the model.*
- *If not, explain why no improvement is possible.*

\vspace{12pt}

\hrule
