## At-Home Exercises

```{r, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

This week, we'll take another look at the Kestilä (2006) results. During this
practical, you will conduct an SEM to replicate the regression analysis of the
Finnish data that you conducted in the [Week 4 In-Class Exercises](in-class-exercises-3.html).

---

###

Load the Finnish subsample of ESS data.

- The relevant data are contained in the [*ess_finland.rds*][fin_data] file.
   - These are the processed Finnish subsample data from the Week 4 exercises.

*Note:* Unless otherwise noted, all the following analyses use these data.

<details>
  <summary>Click to show code</summary>
  
```{r, echo = FALSE}
dataDir <- "../../../../data/"
codeDir <- "../../code/"

source(paste0(codeDir, "supportFunctions.R"))

ess_fin <- readRDS(paste0(dataDir, "ess_finland.rds"))
```

```{r, eval = FALSE}
ess_fin <- readRDS("ess_finland.rds")
```

</details>

---

We need to do a little data processing before we can fit the regression model.
At the moment, `lavaan` will not automatically convert a factor variable into
dummy codes. So, we need to create explicit dummy codes for the two factors
we'll use as predictors in our regression analysis: *sex*  and *political
orientation*.  

---

###

Convert the *sex* and *political interest* factors into dummy codes.

<details>
  <summary>Click to show code</summary>
  
```{r}
library(dplyr)

## Create a dummy codes by broadcasting a logical test on the factor levels:
ess_fin <- mutate(ess_fin,
                  female = ifelse(sex == "Female", 1, 0),
                  hi_pol_interest = ifelse(polintr_bin == "High Interest", 1, 0)
                  )

## Check the results:
with(ess_fin, table(dummy = female, factor = sex))
with(ess_fin, table(dummy = hi_pol_interest, factor = polintr_bin))
```
  <details>
    <summary>Click for explanation</summary>

In R, we have several ways of converting a factor into an appropriate set of 
dummy codes.

- We could use the `dplyr::recode()` function as we did last week.
- We can use the `model.matrix()` function to define a design matrix based on 
the inherent contrast attribute of the factor.
    - Missing data will cause problems here.
- We can us `as.numeric()` to revert the factor to its underlying numeric 
representation {Male = 1, Female = 2} and use arithmetic to convert {1, 2} 
$\rightarrow$ {0, 1}.

When our factor only has two levels, though, the `ifelse()` function is the 
simplest way.

  </details>
</details>

---

We are now ready to estimate our latent regression model. Specifically, we want
to combine the three OLS regression models that you ran in \@ref(efaRegression)
into a single SEM that we will estimate in `lavaan`.

The following path diagram shows the intended theoretical model.

<!------------------------------------------------------------------------------
\[
Y_{trust\_inst} = \beta_1 X_{age} + \beta_2 X_{female} + \beta_3 X_{edu\_years} +
                  \beta_4 X_{hi\_pol\_interest} + \beta_5 X_{lrscale} + 
                  \varepsilon_Y
\]
------------------------------------------------------------------------------->

```{r echo = FALSE}
library(semPlot)

mod_sem <- '
## Define the latent DVs:
institutions =~ trstlgl + trstplc + trstun + trstep + trstprl
satisfaction =~ stfhlth + stfedu  + stfeco + stfgov + stfdem
politicians  =~ pltinvt + pltcare + trstplt

## Specify the structural relations:
institutions +
satisfaction +
politicians ~ female + age + eduyrs + hi_pol_interest + lrscale
'

sem(mod_sem, data = ess_fin, fixed.x = FALSE) %>%
  semPaths(layout = "tree2",
           rotation = 2,
           residuals = FALSE,
           sizeMan = 10,
           sizeMan2 = 3,
           sizeLat = 15,
           sizeLat2 = 8,
           curve = 2,
           esize = 1,
           asize = 2,
           cardinal = "manifest",
           nCharNodes = 0)
```

Although the variances are not included in this path diagram, all variables in
the model (including the observed predictor variables) are random.

---

###

Define the `lavaan` model syntax for the SEM shown above.

- Use the definition of the `institutions`, `satsifaction`, and `politicians`
  factors from \@ref(cfaSyntax) to define the DVs.
- Covary the three latent factors.
- Covary the five predictors.

<details>
<summary>Click to show code</summary>

```{r}
mod_sem <- '
## Define the latent DVs:
institutions =~ trstlgl + trstplc + trstun + trstep + trstprl
satisfaction =~ stfhlth + stfedu  + stfeco + stfgov + stfdem
politicians  =~ pltinvt + pltcare + trstplt

## Specify the structural relations:
institutions ~ female + age + eduyrs + hi_pol_interest + lrscale
satisfaction ~ female + age + eduyrs + hi_pol_interest + lrscale
politicians  ~ female + age + eduyrs + hi_pol_interest + lrscale
'
```

<details>
  <summary>Click for explanation</summary>

- We simply need to add a line defining the latent regression paths to our old
  CFA syntax.
- We don't need to specify the covariances in the syntax. We can use options in
  the `sem()` function to request those estimates.

  </details>
</details>

---

###

Estimate the SEM, and summarize the results.

- Fit the model to the processed Finnish subsample from above.
- Estimate the model using `lavaan::sem()`.
- Request the standardized parameter estimates with the summary.
- Request the $R^2$ estimates with the summary.

<details>
  <summary>Click to show code</summary>

```{r}
library(lavaan)

## Fit the SEM:
fit_sem <- sem(mod_sem, data = ess_fin, fixed.x = FALSE)

## Summarize the results:
summary(fit_sem, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
```

<details>
  <summary>Click for explanation</summary>

The `fixed.x = FALSE` argument tells **lavaan** to model the predictors as random
variables. By default, **lavaan** will covary any random predictor variables. So,
we don't need to make any other changes to the usual procedure.

  </details>
</details>

---

###

Finally, we will rerun the latent regression model from above as a path model 
with the factor scores from \@ref(factorScores) acting as the DVs.

- Rerun the above SEM as a path model wherein the EFA-derived *Trust in 
  Institutions*, *Satisfaction with Political Systems*, and *Trust in Politicians*
  factor scores act as the DVs.
   - Request the standardized parameter estimates with the summary.
   - Request the $R^2$ estimates with the summary.

<details>
  <summary>Click to show code</summary>

```{r}
## Define the model syntax for the path analysis:
mod_pa <- '
trust_inst ~ female + age + eduyrs + hi_pol_interest + lrscale
satisfy    ~ female + age + eduyrs + hi_pol_interest + lrscale
trust_pol  ~ female + age + eduyrs + hi_pol_interest + lrscale
'

## Estimate the path model:
fit_pa <- sem(mod_pa, data = ess_fin, fixed.x = FALSE)

## Summarize the results:
summary(fit_pa, standardized = TRUE, rsquare = TRUE)
```

  <details>
    <summary>Click to show explanation</summary>

We don't so anything particularly special here. We simply rerun our latent 
regression as a path analysis with the EFA-derived factor scores as the DVs.

  </details>
</details>

---

###

Compare the results from the path analysis to the SEM-based results.

- Does it matter whether we use a latent variable or a factor score to define 
the DV?

*Hint:* When comparing parameter estimates, use the fully standardized estimates 
(i.e., the values in the column labeled `Std.all`).

<details>
  <summary>Click to show code</summary>

*Note:* The "supportFunction.R" script that we source below isn't a necessary
part of the solution. This script defines a bunch of convenience functions. One
of these functions, `partSummary()`, allows us to print selected pieces of the 
model summary.

```{r, eval = FALSE}
## Source a script of convenience function definitions:
source("supportFunctions.R")
```

```{r}
## View the regression estimates from the SEM:
partSummary(fit_sem, 8, standardized = TRUE)

## View the regression estimates from the path analysis:
partSummary(fit_pa, 7, standardized = TRUE)

## View the R-squared estimates from the SEM:
partSummary(fit_sem, 11, rsquare = TRUE)

## View the R-squared estimates from the SEM:
partSummary(fit_pa, 10, rsquare = TRUE)
```

```{r, eval = FALSE, echo = FALSE}
r2_sem <- summary(fit_sem, rsquare = TRUE)$pe %>% 
  filter(op == "r2", rhs == "institutions") %>% 
  select(est) %>% 
  as.numeric()
r2_pa <- inspect(fit_pa, "r2")
```

  <details>
    <summary>Click for explanation</summary>

It certainly looks like the way we define the DV has a meaningful impact. The 
patterns of significance differ between the two sets of regression slopes, and
the $R^2$ values are larger for the *Institutions* and *Satisfaction* factors in
the SEM, and the $R^2$ for the *Politicians* factor is higher in the path analysis.

  </details>
</details>

---

End of At-Home Exercises

---

[fin_data]: https://surfdrive.surf.nl/files/index.php/s/rN8FuscpJTLsB4h/download
