## At-Home Exercises

In these exercise we will analyze the *suicide_risk.rds* dataset. These analyses 
are based on the following paper.

[
Metha, A., Chen, E, Mulvenon, S., and Dode, I. (1998). A theoretical model of 
suicide risk. *Archives of Suicide Research, 4*, 115--133.
](https://doi.org/10.1080/13811119808260442)

The *suicide_risk.rds* dataset was simulated from the covariance matrices 
presented on p.123 of the above paper. These data contain the following variables.

1. `suirisk`: A numeric variable representing suicide risk
1. `subabuse`: A numeric variable representing severity of substance abuse
1. `hopeless`: A numeric variable representing degree of hopelessness
1. `selfesteem`: A numeric variable representing levels of self-esteem
1. `depression`: A numeric variable representing depression levels
1. `sex`: A factor recording biological sex

We will use these data to investigate if *sex* moderates the effects of
*substance abuse*, *hopelessness*, *self-esteem*, and *depression* on *suicide
risk*.

---

###

Load the *suicide_risk.rds* dataset.

<details>
  <summary>Click for explanation</summary>

```{r, echo = FALSE}
suicide <- readRDS(paste0(dataDir, "suicide_risk.rds"))
```

```{r, eval = FALSE}
suicide <- readRDS("suicide_risk.rds")
```

</details>

---

###

Use `ggplot()` to visualize the bivariate relations between the four continuous 
predictors and *suicide risk*.

- Based on these figures, do you think a linear relation between *suicide risk*
and the four continuous predictors is reasonable?

<details>
  <summary>Click for explanation</summary>

```{r}
library(ggplot2)

## Create the basic plots for each predictor:
p1 <- ggplot(suicide, aes(x = subabuse, y = suirisk)) 
p2 <- ggplot(suicide, aes(x = hopeless, y = suirisk)) 
p3 <- ggplot(suicide, aes(x = selfesteem, y = suirisk)) 
p4 <- ggplot(suicide, aes(x = depression, y = suirisk)) 

## Add the details and print:
p1 +
  geom_point() +  # Add the scatterplot points
  geom_smooth() + # Add a smoothed trend
  theme_classic() # Use a nicer theme

p2 + geom_point() + geom_smooth() + theme_classic()
p3 + geom_point() + geom_smooth() + theme_classic()
p4 + geom_point() + geom_smooth() + theme_classic()
```

We can also use the `grid.arrange()` function from the **gridExtra** package to
join these four plots into a single figure.

```{r}
library(gridExtra)

grid.arrange(p1 + geom_point() + geom_smooth() + theme_classic(),
             p2 + geom_point() + geom_smooth() + theme_classic(),
             p3 + geom_point() + geom_smooth() + theme_classic(),
             p4 + geom_point() + geom_smooth() + theme_classic(), 
             ncol = 2)
```

There does appear to be an association between each of the predictors and 
*suicide risk*. The trends don't look perfectly linear, but they don't look too 
systematically nonlinear either, so modeling these relationships as linear seems
sensible.

</details>

---
 	 
###

Modify the scatterplots from above by coloring the points and trend lines 
according to `sex`.

- Based on these figures, do you think *sex* may moderate (some of) these 
bivariate relations?

<details>
  <summary>Click for explanation</summary>

In the `ggplot()` call, we simply need to define assign `sex` to the "colour" 
aesthetic.

```{r}
## Create the basic plots for each predictor:
p1 <- ggplot(suicide, aes(x = subabuse, y = suirisk, colour = sex)) 
p2 <- ggplot(suicide, aes(x = hopeless, y = suirisk, colour = sex)) 
p3 <- ggplot(suicide, aes(x = selfesteem, y = suirisk, colour = sex)) 
p4 <- ggplot(suicide, aes(x = depression, y = suirisk, colour = sex))

grid.arrange(p1 + geom_point() + geom_smooth() + theme_classic(),
             p2 + geom_point() + geom_smooth() + theme_classic(),
             p3 + geom_point() + geom_smooth() + theme_classic(),
             p4 + geom_point() + geom_smooth() + theme_classic(), 
             ncol = 2)
```

Of all the predictors, it looks like *substance abuse* shows the greatest 
differences between the male and female trends. The other three predictors 
demonstrate pretty similar effects for both sexes.

</details>

---

To test moderation, we need to include interactions (i.e., product terms) 
between the moderator variable(s) and the focal predictor(s) on the RHS of our 
regression equation. If we're using the `lm()` function to estimate our models 
via OLS regression, we need only specify the appropriate product terms in the 
model syntax. 

---

### {#olsMod}

Use the `lm()` function to test the moderating influence of `sex` on the four 
focal effects with OLS regression.

- Does *sex* significantly moderate the four focal effects?

*Hint:* To specify an interaction, simply "multiply" the variable names for the
moderator and focal predictor using the `*` symbol in the formula syntax.

- E.g., `y ~ x * z`

<details>
  <summary>Click for explanation</summary>

```{r}
out <- lm(suirisk ~ sex * subabuse + sex * hopeless + sex * selfesteem + sex * depression, 
          data = suicide)

summary(out)
```

```{r, echo = FALSE}
tmp <- summary(out)$coef["sexmale:depression", ]

b <- tmp["Estimate"] 
t <- tmp["t value"]
p <- tmp["Pr(>|t|)"]

df <- out$df.residual
```

It doesn't look like *sex* has too much affect. Only the effect of *depression* 
on *suicide risk* was significantly moderated by *sex* ($\beta = `r round(b, 3)`$, 
$t[`r df`] = `r round(t, 2)`$, $p = `r round(p, 3)`$)

</details>

---

Of course, we can do an analogous analysis via path modeling in **lavaan**. 
Unfortunately, **lavaan** will not (yet) dummy code our factors, so we need to
manually create a dummy coded version of *sex*.

---

###

Dummy code the `sex` factor.

<details>
  <summary>Click for explanation</summary>

You have seen several ways of doing this in previous practicals. Here, I'll use
the quick-and-dirty `ifelse()` approach.

```{r}
suicide$female <- ifelse(suicide$sex == "female", 1, 0)
```

</details>

---

###

Define the **lavaan** model syntax for the same moderated regression model you 
estimated in \@ref(olsMod).

*Hints:*

- To define an interaction term in lavaan syntax, use the `:` operator instead 
of `*`.
- You will need to explicitly include the variable names for the individual 
predictors (as well as the interaction terms) in the model syntax.

<details>
  <summary>Click for explanation</summary>
  
```{r}
mod <- '
suirisk ~ subabuse + hopeless + selfesteem + depression + female
suirisk ~ female:subabuse + female:hopeless + female:selfesteem + female:depression
'
```

Note that I've broken the formula definition into two lines. Doing so has no
effect on the model's specification or estimation but keeps the syntax a bit 
tidier (by keeping the lines from getting too long).

</details>

---

### {#pathMod}

Use `lavaan::sem()` to estimate the above model.

- What are your conclusions vis-à-vis the moderating influence of *sex*?
- Do these results differ from those of the OLS regression?
- What proportion of the variability in *suicide risk* is explained by this model?

<details>
  <summary>Click for explanation</summary>

```{r}
library(lavaan)

out <- sem(mod, data = suicide)
summary(out, rsquare = TRUE)
```

- The path analysis gives the same results as the OLS regression. So, the 
conclusions are also the same.
- The model explains `r (100 * lavInspect(out, "r2")) %>% round(2)`% 
of the variability in *suicide risk*.

</details>

---

Based on the results above, it looks like *sex* does not have much moderating
influence. If we are trying to build a statistical model of *suicide risk*, we 
may question if the overall moderating effect of *sex* really improves the model.

We may also want to evaluate the significance of the focal effects, themselves. 
In the moderated model, the estimates for the focal predictors are *conditional 
effects*, not ordinary partial regression weights. In this particular case, these
conditional effects represent the effect of each focal predictor on the outcome
for males (i.e., the reference group of the moderator).

If we simply want to know if these focal predictors significantly affect *suicide
risk*, without conditioning on *sex*, we need to estimate a restricted model
containing no moderation (i.e., the additive model with all the same predictors 
but no interactions).

---

###

Specify the **lavaan** model syntax for the restricted model described above.

<details>
  <summary>Click for explanation</summary>
  
```{r}
mod <- '
suirisk ~ subabuse + hopeless + selfesteem + depression + female
'
```

</details>

---

###

Use `lavaan::sem()` to estimate the above model.

- Are the focal predictors significantly associated with *suicide risk*?
- Is *sex* significantly associated with *suicide risk*?
- How much variability in *suicide risk* is explained by these predictors?

<details>
  <summary>Click for explanation</summary>

```{r}
out_res <- sem(mod, data = suicide)
summary(out_res, rsquare = TRUE)
```

```{r, echo = FALSE}
tmp <- summary(out_res)$pe %>% 
  filter(op == "~", rhs %in% c("subabuse", "depression", "female"))

b1 <- tmp[1, "est"] 
z1 <- tmp[1, "z"]

b2 <- tmp[2, "est"]
z2 <- tmp[2, "z"]

b3 <- tmp[3, "est"]
z3 <- tmp[3, "z"]
p3 <- tmp[3, "pvalue"]
```

- Only *substance abuse* ($\beta = `r round(b1, 3)`$, $z = `r round(z1, 2)`$, 
$p < 0.001$) and *depression* ($\beta = `r round(b2, 3)`$, $z = `r round(z2, 2)`$, 
$p < 0.001$) are significant predictors of *suicide risk*.
- No, *sex* is not a significant predictor of *suicide risk* 
($\beta = `r round(b3, 3)`$, $z = `r round(z3, 2)`$, $p = `r round(p3, 3)`$).
- These five predictors explain `r (100 * lavInspect(out_res, "r2")) %>% round(2)`% 
of the variability in *suicide risk*.

</details>

---

###

Compare the full model (including interactions) and the restricted model (without
interactions).

- How much additional variability in *suicide risk* do we explain by including the 
four interaction terms?
- Can we conduct a $\Delta \chi^2$ test to compare these two models?
    - If you think so, conduct the test and interpret the results.
    - If you do not think so, explain why not.

<details>
  <summary>Click for explanation</summary>

```{r, echo = FALSE}
r2Diff <- lavInspect(out, "r2") - lavInspect(out_res, "r2")
```

To see how much additional variability we explain, we simply subtract the $R^2$
for the restricted model from the $R^2$ for the full model. 

- Adding the interaction terms explains `r round(100 * r2Diff, 2)`% more 
variability in *suicide risk*.

Unfortunately, we cannot use a $\Delta \chi^2$ test to compare these models 
because they are both saturated (hence, they both have perfect fit). Also, these
models are not nested because we added four new variables (hence, added four new 
columns and rows to the covariance matrix) when adding the interaction terms.

- If we want, we can manually implement a $\Delta R^2$ test analogous to what 
we would do in OLS regression, but doing so is a bit of a faff.

```{r}
## Extract the residual variance:
eF <- lavInspect(out, "theta")["suirisk", "suirisk"]
eR <- lavInspect(out_res, "theta")["suirisk", "suirisk"]

## Compute the residual DF:
n   <- lavInspect(out, "nobs")
dfF <- n - fitMeasures(out, "npar")
dfR <- n - fitMeasures(out_res, "npar")

## Compute the F statistic for the R^2 difference:
(f <- ((eR - eF) / (dfR - dfF)) / (eF / dfF))

## Compute the p-value by comparing to an F-distribution
(p <- pf(f, dfR - dfF, dfF, lower.tail = FALSE))
```

The F-statistic for the $\Delta R^2$ test is significant 
($\Delta R^2 = `r round(r2Diff, 3)`$, 
$F[`r dfR - dfF`, `r dfF`] = `r round(f, 2)`$,
$p = `r round(p, 3)`$), so we can say that including the moderating influence of 
*sex* explains significantly more variability in *suicide risk*. Of course, you 
could certainly question if an additional `r round(100 * r2Diff, 2)`% of 
variance explained is a meaningful increase. Regardless of how you judge the 
increase in explained variance, though, we cannot necessarily say that adding 
the interaction terms produces a "better" model. Both models perfectly recreate 
the observed data; they are statistically equivalent representations of the 
phenomenon.

</details>

---

End of At-Home Exercises

---
