---
title: "Group 42"
author: ["Zaphod Beeblebrox", "Tricia McMillan", "Marvin T.P. Andy"]
date: "1970-01-01"
format:
  pdf:
    number-sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lavaan)
library(knitr)
```

# Theory

## Which dataset are you using?

## Describe your hypothesized measurement model.

## Provide a path diagram of your theoretical measurement model.

\vspace{12pt}

# Estimation

## Provide the lavaan syntax for your model.

```{r}

```

## Use the `lavaan::cfa()` function to estimate your model.

```{r}

```

\vspace{12pt}

# Results

## Report the relevant parameter estimates from your fitted model.

```{r, echo = FALSE}

```

## Report the $R^2$ estimates from your fitted model.

```{r, echo = FALSE}

```

## Report the model fit indices for your estimated model.

```{r, echo = FALSE}

```

\vspace{12pt}

# Interpretation

## Do your results support the validity of your hypothesized model?

- *Explain why or why not.*

## Do you think the model could be improved?

- *If so, explain how you would modify the model.*
- *If not, explain why no improvement is possible.*

---
